---
title: "Barplots"
author: "Albert Y. Kim"
date: "Mon Oct 10, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(babynames)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
# Bar Chart
babynames_counts <- babynames %>% 
  mutate(decade = floor(year / 10) * 10) %>% 
  filter(name=="Hayden" | name == "Carlos" | name == "Ethan") %>% 
  filter(decade >= 1990)
```


## 5NG

Today:

1. Scatterplot AKA bivariate plot
1. Linegraph
1. Boxplot
1. Histogram
1. **Barplot AKA barchart AKA bargraph**





## Barplots

> * In theory: Barplots display information about a single **categorical** variable as the `x` aesthetic.
> * In practice: We'll see shortly there is a twist.
> * The y-axis displays notions of **relative frequency** i.e. which values occur more than 
others. This is (in theory) not a variable in the data set, but rather is computed.





## Barplots

> * `geom_bar()` is the trickiest of the 5NG in `ggplot2`, hence me presenting it last
> * Correspondingly, we'll use it in limited capacity in this class
> * Also take me at my word for certain commands.
> * If, however, you are feeling adventurous, feel free to ask me anyway!





## Barplots

Recall from Lec05 Slide 16, we displayed

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
# Bar Chart
babynames_counts %>% 
  group_by(name) %>% 
  summarise(n=sum(n)) %>% 
  ggplot(data=., aes(x=name, y=n)) +
  geom_bar(stat="identity") +
  labs(x="Name", y="Count", title="Name Counts since 1990")
```




## Input Data

The input `data` to the `ggplot()` for `geom_bar()` has to be **pre-counted** per value of the categorical variable.






## Input Data

So while the raw `babynames` data looks like...

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
babynames_counts %>% 
  select(-prop, -decade) %>% 
  kable()
```




## Input Data

The input data needs to be **pre-counted** for each name.  Let's call this data
frame `babynames_counted`.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
babynames_counts %>% 
  select(-prop, -decade) %>% 
  group_by(name) %>% 
  summarise(n=sum(n)) %>% 
  kable()
```

  

## ggplot Command

* We assign the `y` `aes()`thetic to be the **pre-counted** variable `n`
* The `stat="identity"` in `geom_bar()`:

```{r, eval=FALSE, echo=TRUE}
ggplot(data=babynames_counted, aes(x=name, y=n)) +
  geom_bar(stat="identity")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
# Bar Chart
babynames_counts %>% 
  group_by(name) %>% 
  summarise(n=sum(n)) %>% 
  ggplot(data=., aes(x=name, y=n)) +
  geom_bar(stat="identity") +
  labs(x="Name", y="Count", title="Name Counts since 1990")
```



