---
title: "Boxplots"
author: "Albert Y. Kim"
date: "Fri Sep 30, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```



## 5NG

Today:

1. Scatterplot AKA bivariate plot
1. Linegraph
1. **Boxplot**
1. Barplot AKA Barchart AKA bargraph
1. Histogram



## Example

If I know your name, I can guess your [age](http://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/).



## Statistics Terminology

* All the plots with either blue/yellow bars red dots are **boxplots** (without whiskers) that indicate the 3 **quartiles**
    + 1st quartile i.e. 25th percentile
    + 2nd quartile i.e. 50th percentile
    + 3rd quartile i.e. 75th percentile



## Statistics Terminology

> * The 2nd quartile is also called the **median**. It is a measure of **center**
> * The width of the bars (3rd quartile - 1st quartile) is the **interquartile range (IQR)**
> * The IQR contains the middle 50% of observations. It is a measure of **spread**
    


## Boxplots

[Breakdown](http://www.datavizcatalogue.com/methods/images/anatomy/box_plot.png) of components.


    
## Why Boxplots?

You can compare **distributions** of values in different groups with a single line.

Look at the chart in this [Planet Money](http://www.npr.org/sections/money/2015/03/19/394057221/how-much-or-little-the-middle-class-makes-in-30-u-s-cities) article. In this case,
you can compare cities with a single vertical line.
