---
title: "Intro to Data Manipulation/Wrangling"
author: "Albert Y. Kim"
date: "Wed Oct 12, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```



## 5MV

Say hello to the 5MV: the **five main verbs**

1. `filter()` rows matching criteria
1. `select()` columns by variable name
1. `mutate()` existing variables to create new ones
1. `arrange()` rows
1. `summarise()` numerical variables that are `group_by()` categorical variables



## 5MV + Joins

Also, we can `_join()` two data frames `by` corresponding variables




## Important Concept: Piping

Piping allows you to take the output of one function and **pipe** it as the input of
the next function. You can string along several pipes to form a single **chain**.

* R Command: `%>%`
* Described as: "_then_". 



## Important Concept: Piping

Ex:  say you want to apply functions `h()` and `g()` and then `f()` on data `x`.  You can do

* `f(g(h(x)))` OR
* `x %>% h() %>% g() %>% f()`

This

> * saves you from confusing nested parentheses
> * emphasizes the **sequential** breaking down of tasks, making it more readable
> * i.e. Do this _then_ do this _then_ do this _then_ ...




## Important Concept: Piping

Pipes are always directed to the first argument of any function. The following three bits of R code
do the same thing, extract all january flights:

```{r,eval=FALSE, echo=TRUE}
library(nycflights13)
data(flights)

# Bit 1: No piping
filter(flights, month == 1)

# Bit 2: Piping. Note no comma
flights %>% filter(month == 1)

# Bit 3: Piping across multiple lines (prefered)
flights %>% 
  filter(month==1)
```
