---
title: "Finishing the 5MV"
author: "Albert Y. Kim"
date: "Fri Oct 21, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```





## Last Time: 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. `mutate()` existing variables to create new ones
1. ~~`arrange()` rows~~





## Today: Finishing 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. ~~`mutate()` existing variables to create new ones~~
1. `arrange()` rows

and `_join()`: backside of cheatsheet, right hand column, top.





## Arrange

Really simple. Either

* `DATASET_NAME %>% arrange(VARIABLE_NAME)` or
* `DATASET_NAME %>% arrange(desc(VARIABLE_NAME))`





## Arrange Example

```{r, eval=FALSE, echo=TRUE}
library(dplyr)

# Create data frame with two variables
test_data <- data_frame(
  name=c("Abbi", "Abbi ", "Ilana", "Ilana", "Ilana"),
  value_1=c(0, 1, 0, 1, 0),
  value_2=c(4, 6, 3, 2, 5)
)

# See contents in console
test_data
```





## Arrange Example

Run these bits of code. Notice the subtle diff between 2 and 3:

```{r, eval=FALSE, echo=TRUE}
# Bit1: Arrange in ascending order of value_1 (the default)
test_data %>% arrange(value_1)

# Bit2: Arrange in descending order of value_1
test_data %>% arrange(desc(value_1))

# Bit3: Arrange in decending order of value_1, and then within
# value_1, arrange in ascending order of value_2
test_data %>% arrange(desc(value_1), value_2)
```





## Combining Data Sets via Joins

And now the last component of data manipulation, and the data portion of this 
class: **joining/merging two data sets**.





## Combining Data Sets via Joins

Imagine we have two data frames `x` and `y`. Run the following:

```{r, eval=FALSE, echo=TRUE}
x <- data_frame(x1=c("A","B","C"), x2=c(1,2,3))
y <- data_frame(x1=c("A","B","D"), x3=c(TRUE,FALSE,TRUE))
x
y
```





## Combining Data Sets via Joins

We join by the `"x1"` variable. Note how it is in quotation marks.

```{r, eval=FALSE, echo=TRUE}
left_join(x, y, by = "x1")
full_join(x, y, by = "x1")
```





## Different Join Variable Name?

What the variable we want to join by has a different name in each data set? 
Here `a` vs `b`:

```{r, eval=FALSE, echo=TRUE}
i <- data_frame(a=c("A","B","C"), x2=c(1,2,3))
j <- data_frame(b=c("A","B","D"), x3=c(TRUE,FALSE,TRUE))
i
j
```





## Different Join Variable Name?

We use the `c()` function.

```{r, eval=FALSE, echo=TRUE}
left_join(i, j, by = c("a" = "b"))
```

It is a bit confusing unfortunately. It's

* `by = c("a"="b")` and not
* `by = c("a"=="b")`





## Different Join Variable Name?

There was an example of such a join in [Chapter 4.7.1](https://rudeboybert.github.io/IntroStatDataSciences/4-viz.html#barplots) on Barplots
where we `join` 

* `flights` data, which had airport codes in `origin`
* `airports` data, which had airport codes in `faa`

```{r, eval=FALSE, echo=TRUE}
inner_join(flights, airports, by = c("origin" = "faa"))
```

This creates a single data set with all information combined.





## Types of Joins

* There are many types of `join` (right-hand column of back of cheatsheet)
* To keep things simple, we'll try to only use:
    + `left_join`
    + `full_join`
* This
[illustration](https://twitter.com/yutannihilation/status/551572539697143808)
succinctly summarizes all of them.


