---
title: "Finishing the 5MV"
author: "Albert Y. Kim"
date: "Fri Oct 21, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```





## Last Time: 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. `mutate()` existing variables to create new ones
1. ~~`arrange()` rows~~






## Today: Finishing 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. ~~`mutate()` existing variables to create new ones~~
1. `arrange()` rows

and `_join()`






## Today: Finishing 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. ~~`mutate()` existing variables to create new ones~~
1. `arrange()` rows

and `_join()`





