---
title: "Histograms"
author: "Albert Y. Kim"
date: "Fri Oct 7, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(okcupiddata)
```



## 5NG

Today:

1. Scatterplot AKA bivariate plot
1. Linegraph
1. Boxplot
1. **Histogram**
1. Barplot AKA Barchart AKA bargraph



## Histograms

> * Midterm: Histograms display information about a single continuous/numerical variable as the `x` aesthetic
> * The y-axis displays notions of **relative frequency** i.e. which values occur more than 
others. This is not a variable in the data set, but rather is computed internally.
> * Yoooooooge definition: they are a visualization of the **statistical distribution**.



## Recall from Lec05 Slide 11

From `okcupiddata` package, the `profiles` data set:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Histogram
profiles %>% 
  ggplot(data=., aes(x=height)) +
  geom_histogram(binwidth = 1) +
  labs(x="height (in inches)", y="count", title="Heights of San Francisco OkCupid Users")
```



## Recall from Lec05 Slide 11

Cleaned up to only have heights between 55 (5'5'') and 80 (6'8'') inches:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Histogram
profiles %>% 
  filter(between(height, 55, 80)) %>% 
  ggplot(data=., aes(x=height)) +
  geom_histogram(binwidth = 1) +
  labs(x="height (in inches)", y="count", title="Heights of San Francisco OkCupid Users")
```












