---
title: "Linegraph"
author: "Albert Y. Kim"
date: "Wed Sep 28, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```



## 5NG

Today:

1. Scatterplot AKA bivariate plot
1. **Linegraph**
1. Boxplot
1. Barplot AKA Barchart AKA bargraph
1. Histogram




## Example

Let's re-run our example from last time:

```{r, eval=FALSE, echo=TRUE}
# Load packages
library(dplyr)
library(ggplot2)

# Create data frame
simple_ex <-
  data_frame(
    A = c(1, 2, 3, 4),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("a", "a", "b", "b")
  )

simple_ex
```

```{r, eval=TRUE, echo=FALSE}
# Load packages
library(dplyr)
library(ggplot2)

# Create data frame
simple_ex <-
  data_frame(
    A = c(1, 2, 3, 4),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("a", "a", "b", "b")
  )
```


## Example 

A statistical graphic is a mapping of `data` variables to `aes()`thetic
attributes of `geom_`etric objects.

```{r, echo=FALSE, fig.height=3}
ggplot(data=simple_ex, aes(x=A, y=B)) + 
  geom_line()
```


## Today

* Drill down on `geom_line()`
* How to have more than one line-graph on the same plot using the `color` aesthetic.
* Introduce two techniques for [**smoothing**](http://blog.revolutionanalytics.com/2010/03/smoothing-time-series-with-r.html). This allows us to separate the "signal" from the "noise"
    + General smoothing
    + Linear regression
    
    




