---
title: "More of the 5MV"
author: "Albert Y. Kim"
date: "Wed Oct 19, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```





## Last Time: 5MV

1. ~~`select()` columns by variable name: **front of cheatsheet, bottom right**~~
1. ~~`filter()` rows matching criteria: **front of cheatsheet, bottom middle.**~~
1. `summarise()` numerical variables that are `group_by()` categorical variables
1. ~~`mutate()` existing variables to create new ones~~
1. ~~`arrange()` rows~~

And piping `%>%` i.e. *then*





## Keyboard Shortcut

Everyone place your cursor in the console and type:

* Mac users: COMMAND+SHIFT+M
* Windows uers: CTRL+SHIFT+M




## Today: 5MV

1. ~~`select()` columns by variable name: front of cheatsheet, bottom right~~
1. ~~`filter()` rows matching criteria: front of cheatsheet, bottom middle~~
1. ~~`summarise()` numerical variables that are `group_by()` categorical variables~~
1. `mutate()` existing variables to create new ones: **back of cheatsheet,
center column** However in this class you probably won't use these often.
1. ~~`arrange()` rows~~





## mutate()

Always of the form

```{r, eval=FALSE, echo=TRUE}
mutate(DATASET_NAME, NEW_VARIABLE_NAME = OLD_VARIABLE_NAMES)
```

or if using piping `%>%` (note no comma)

```{r, eval=FALSE, echo=TRUE}
DATASET_NAME %>% mutate(NEW_VARIABLE_NAME = OLD_VARIABLE_NAMES)
```





## Example

Load the following in your console

```{r, eval=FALSE, echo=TRUE}
library(dplyr)

# Create data frame with two variables
test_data <- data_frame(
  name=c("Kimmy", "Kimmy", "Kimmy", "Titus", "Titus"),
  value=c(1, 2, 3, 4, 5)
)

# See contents in console
test_data
```





## Mutate

Test out the following bits individually in your console:

```{r, eval=FALSE, echo=TRUE}
# Bit1: Mutate the variable value by adding 200, overwrite value
mutate(test_data, value = value+200)

# Bit2: Same thing, but with piping. Note there is no comma now 
# as the %>% by default # pipes to the first spot
test_data %>% mutate(value = value+200)

# Bit3: Mutate the variable value by adding 200, but now save
# in a new variable called new_value
test_data %>% mutate(new_value = value+200)
```




