---
title: "More Grammar of Graphics"
author: "Albert Y. Kim"
date: "Fri Sep 23, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```



## Data Visualization

> * We are building up to doing data visualization in R via the `ggplot2` package
> * Last time we reverse-engineered the grammar from graphical outputs
> * Today we (forward) engineer them




## Refresher: The Grammar of Graphics

A statistical graphic is a mapping of `data` variables to `aes()`thetic
attributes of `geom_`etric objects.



## Refresher: 5NG

The **five named graphs** we'll see in this class

1. Scatterplot AKA bivariate plot
1. Line-graph
1. Boxplot
1. Barplot AKA Barchart AKA bargraph
1. Histogram



## Data

Consider the following data in tidy format:

```{r}
simple <-
  data_frame(
    A = c(1, 2, 3, 4),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("a", "a", "b", "b")
  )
simple %>% kable(format="markdown")
```


## Learning Check

Draw the 5 graphics below, where the `x`-axis is variable `A`, the `y`-axis is variable `B`, and

1. A scatter plot
1. A scatter plot where the `color` of the points corresponds to `D`
1. A scatter plot where the `size` of the points corresponds to `C`
1. A line graph
1. A line graph where the `color` of the line corresponds to `D`






