---
title: "Learning Checks"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: cosmo
    highlight: tango
    df_print: paged
---

<style>
h1{font-weight: bold;}
h2{color: #3399ff;}
h3{color: #3399ff;}
slides > slide.backdrop {background: white;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!----------------------------------------------------------------------------->
# Lec04 - Wed 2/22: 5NG

Based on the 5NG examples in today's slides

* **Learning Check 1**: Following the example of Napoleon's march, identify the elements of the Grammar of Graphics:
    1. identify the `data` variables being displayed and what type of variable they are
    1. identify the `aes()`thetic attribute of the `geom_`etric object the above `data` variables are being mapped to
* **Learning Check 2**: Answer the following questions:
    1. Does spending more on a movie yield higher IMDB ratings?
    1. Why are there drops in the number of flights?
    1. Click <a target="_blank" class="page-link" href="http://www.datavizcatalogue.com/methods/images/anatomy/box_plot.png">here</a> for an explanation of boxplots. About what proportion of manual car models sold between 1984 and 2015 got 20 mpg or worse mileage?
    1. About how many babies were named "Hayden" between 1990-2014?
    1. What are the smallest and largest visible heights and what do you think of them? Also, think of one graph improvement to better convey information about SF OkCupid users.
















<!----------------------------------------------------------------------------->
# Lec03 - Mon 2/20: Tidy Data

Do the 16 Learning Checks in <a target="_blank" class="page-link" href="https://ismayc.github.io/moderndiver-book/3-tidy.html">Chapter 3 of ModernDive: Tidy Data</a>. You do not need to submit these.

## Discussion

**3.1** 3.2 is an example!

**3.2** Since there are three variable at play (Date, Price, Stock Name), there should be three columns!

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
data.frame(
  Date = as.Date('2009-01-01') + 0:4,
  Boeing = paste("$", c("173.55", "172.61", "173.86", "170.77", "174.29"), sep=""),
  Amazon = paste("$", c("174.90", "171.42", "171.58", "173.89", "170.16"), sep=""),
  Google = paste("$", c("174.34", "170.04", "173.65", "174.87", "172.19") ,sep="")
) %>% 
  gather(`Stock Name`, Price, -Date) %>% 
  knitr::kable()
```

**3.3 What does any ONE row in thd `flights` dataset refer to?**  Data on a flight. Not a flight path! Example: 

* a flight path would be United 1545 to Houston
* a flight would be United 1545 to Houston *2013/1/1 at 5:15am*


**3.4 What are some examples in this dataset of categorical variables? What makes them different than quantitative variables?**

Hint: Type `?flights` in the console to see what all the variables mean!

* Cateogorical:
    + `carrier` the company
    + `dest` the destination
    + `flight` the flight number. Even though this is a number, its simply a label. Example United 1545 isn't "less than" United 1714
* Quantitative:
    + `distance` the distance in miles
    + `time_hour` time

**3.5 What does `int`, `dbl`, and `chr` mean in the output above?** 

* `int`: integer. Used to count things i.e. a discrete value. Ex: the # of cars parked in a lot
* `dbl`: double. Used to measure things. i.e. a continuous value. Ex: your height in inches
* `chr`: character. i.e. text


**3.6 & 3.7** 19 columns (variables) and 336,776 rows (observations i.e. flights)

**3.8** `weather`, `planes`, `airports`, `airlines` data sets.

The observational units, i.e. what each row corresponds to:

* `weather`: weather at a given `origin` (EWR, JFK, LGA) for a given hour i.e. `year`, `month`, `day`, `hour`
* `planes`: a physical aircraft
* `airports`: an airport in the US
* `airlines`: an airline company

**3.9** See `?airports` help file

**3.10** Identification Variables

* In the `weather` example in LC3.8, the combination of `origin`, `year`, `month`, `day`, `hour`  are identification variables as they identify the observation in question.
* Anything else pertains to observations: `temp`, `humid`, `wind_speed`, etc.

**3.11 What are common characteristics of “tidy” datasets?** Described in Lecture slides.

**3.12 What makes “tidy” datasets useful for organizing data?** Organized way of viewing data. We'll see later that this format is required for the `ggplot2` and `dplyr` packages for data visualization and manipulation.

**3.13** There are 2 variables below, but what does each row correspond to? We don't know b/c there are no identification variables.

```{r, echo=FALSE, message=FALSE}
students <- c(4, 6)
faculty <- c(2, 3)
data_frame("students" = students, "faculty" = faculty) %>% 
  knitr::kable()
```

**3.14** We need at least a third variable to identify the observations. For example a variable "Department".

**3.15** Sociology example

* Each row is a member of a university.
* Variables are the columns
* `TRUE` and `FALSE`. This is called a logical variable AKA a boolean variable. `1` and `0` can also be used

**3.16** We can easily `_join` them with other data sets! For example, can we join the `flights` data with the `planes` data?  We'll see this more in Chapter 4!



****









<!----------------------------------------------------------------------------->
# Lec02 - Thu 2/16: R Packages

You will be getting your first experience with:

* RStudio: i.e. The Dashboard
* R Packages: Extending the functionality
* `View()`: A command for viewing your data
* R Scripts: Text files to save your work
* Answering questions with data. In this case: baby name popularity


## Setting Up

As described in today's lecture slides: In RStudio (not DataCamp):

1. Install the `ggplot2`, `dplyr`, and `babynames` packages.
1. Load the `ggplot2`, `dplyr`, and `babynames` packages.


## `View()` Your Data

Load the `babynames` dataset in the RStudio viewer by running the following in 
the console. You should get in the habit of **always** `View()`ing your data first!

```{r, eval=FALSE}
View(babynames)
```

* Scroll through the viewer to get an overall feel for your data:
* Filter your data in the viewer
    + Click on the **Filter** button
    + Click in the white boxes under the variable names and
        + Play with any sliders for numerical variables: year, n, prop
        + Enter in values to view subsets of rows: sex, name


## R Scripts

There are two ways to run commands in the R console: Either

1. Typing them directly in the console and pressing enter (as you just did).
1. Saving them in a `.R` *R Script* and passing them to the console.

Do the following:

* In RStudio menu bar -> File -> New File -> R Script
* Save the file as `babynames`. Note:
    + A file extension `.R` gets added: `babynames.R`
    + You should see `babynames.R` in the File panel of RStudio:
* Cut/paste the contents of the grey block below into `babynames.R` and save it again.
* Run all the code in the console by highlighting it and pressing
    + Mac users: COMMAND+ENTER
    + PC users: CTRL+ENTER
* Comment on what you see


```{r, eval=FALSE}
# This bit of code loads packages:
library(ggplot2)
library(dplyr)
library(babynames)

# The code below shows the popularity trend for the name "Jeffrey" amongst 
# males. You do not need to understand this code for now; Just run it.
baby_name <- "Jeffrey"
baby_sex <- "M"

single_name <- babynames %>% 
  filter(name==baby_name & sex==baby_sex)

ggplot(data=single_name, aes(x=year, y=prop)) + 
  geom_line() +
  xlim(c(1880, 2014)) +
  ylim(c(0, NA)) +
  labs(x="Year", y="Proportion", title = paste("Prop of", baby_sex, "Named", baby_name))
```


## Today's Exercise

Investigate your hypothesized names that are "modern", "old-fashioned", and "back in vogue" by

* Changing **only** what gets assigned to
    + `baby_name`
    + `baby_sex`
* Running the appropriate lines of code in the console.



