---
title: "Learning Checks"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: cosmo
    highlight: tango
    df_print: paged
---

<style>
h1{font-weight: bold;}
h2{color: #3399ff;}
h3{color: #3399ff;}
slides > slide.backdrop {background: white;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!----------------------------------------------------------------------------->
# Lec03 - Thu 2/16: Intro to Tidy Data

Do the 16 Learning Checks in <a target="_blank" class="page-link" href="https://ismayc.github.io/moderndiver-book/3-tidy.html">Chapter 3 of ModernDive: Tidy Data</a>. You do not need to submit these.








<!----------------------------------------------------------------------------->
# Lec02 - Thu 2/16: R Packages

You will be getting your first experience with:

* RStudio: i.e. The Dashboard
* R Packages: Extending the functionality
* `View()`: A command for viewing your data
* R Scripts: Text files to save your work
* Answering questions with data. In this case: baby name popularity


## Setting Up

As described in today's lecture slides: In RStudio (not DataCamp):

1. Install the `ggplot2`, `dplyr`, and `babynames` packages.
1. Load the `ggplot2`, `dplyr`, and `babynames` packages.


## `View()` Your Data

Load the `babynames` dataset in the RStudio viewer by running the following in 
the console. You should get in the habit of **always** `View()`ing your data first!

```{r, eval=FALSE}
View(babynames)
```

* Scroll through the viewer to get an overall feel for your data:
* Filter your data in the viewer
    + Click on the **Filter** button
    + Click in the white boxes under the variable names and
        + Play with any sliders for numerical variables: year, n, prop
        + Enter in values to view subsets of rows: sex, name


## R Scripts

There are two ways to run commands in the R console: Either

1. Typing them directly in the console and pressing enter (as you just did).
1. Saving them in a `.R` *R Script* and passing them to the console.

Do the following:

* In RStudio menu bar -> File -> New File -> R Script
* Save the file as `babynames`. Note:
    + A file extension `.R` gets added: `babynames.R`
    + You should see `babynames.R` in the File panel of RStudio:
* Cut/paste the contents of the grey block below into `babynames.R` and save it again.
* Run all the code in the console by highlighting it and pressing
    + Mac users: COMMAND+ENTER
    + PC users: CTRL+ENTER
* Comment on what you see


```{r, eval=FALSE}
# This bit of code loads packages:
library(ggplot2)
library(dplyr)
library(babynames)

# The code below shows the popularity trend for the name "Jeffrey" amongst 
# males. You do not need to understand this code for now; Just run it.
baby_name <- "Jeffrey"
baby_sex <- "M"

single_name <- babynames %>% 
  filter(name==baby_name & sex==baby_sex)

ggplot(data=single_name, aes(x=year, y=prop)) + 
  geom_line() +
  xlim(c(1880, 2014)) +
  ylim(c(0, NA)) +
  labs(x="Year", y="Proportion", title = paste("Prop of", baby_sex, "Named", baby_name))
```


## Today's Exercise

Investigate your hypothesized names that are "modern", "old-fashioned", and "back in vogue" by

* Changing **only** what gets assigned to
    + `baby_name`
    + `baby_sex`
* Running the appropriate lines of code in the console.



