---
title: "Designing Experiments"
author: "Albert Y. Kim"
date: "Fri Nov 4, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(mosaic)
```



## Recall

* Say its the early 1900's, and you are a statistician and you meet someone who claims to be able to tell by tasting whether the tea or the milk was added first to a cup.
* You call BS and think they are just guessing.
* Say you have 8 cups, tea, and milk handy. How would you **design an experiment** to test whether a) they can really tell which came first or b) they are just guessing?
* Brainstorm all the components of this experiment with your seatmates, then think about how you can implement this with `resampl()`ing.



## Picking Pods at Random

```{r, message=FALSE, eval=FALSE, echo=FALSE}

```


