---
title: "Hypothesis Testing Part III"
author: "Albert Y. Kim"
date: "Mon Nov 16, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(mosaic)
```




## Today

More answering questions with data.




## Concept 1: EDA

> * Before using any statistical tool, always do an **exploratory data analysis (EDA)** to look at your data.
> * You might not even need statistics.




## Concept 2: Wrapper Function

> * Wikipedia: A **wrapper function** is computer code whose main purpose is to
call more complex computer code with little or no additional work.
> * The wrapper exists simply to "hide" more complex code and/or to simplify its use.




## Concept 3: Setting Seed Value

A computer can't generate pure random numbers, but **pseudorandom** results:

> * Takes a seed value and returns numbers that are indistinguishable from true randomness
> * If you don't give a seed, the computer picks one based on how much disk space + clock
> * If you do give the same seed, you'll get **replicable** pseudorandom results.




## Run the Following in Console

```{r, eval=FALSE}
library(mosaic)
basket <- c("apple", "orange", "mango", "pear")

# Seed is set automatically by computer
shuffle(basket)

# Set seed
set.seed(76)
shuffle(basket)

# Set seed to same value to get replicable pseudorandom results.
set.seed(76)
shuffle(basket)
```




## Setting Seed Value

My favorite seed value is [76](http://www.macleans.ca/wp-content/uploads/2014/05/MAC21_SUBBAN_CAROUSEL.jpg).







