---
title: "p-Value"
author: "Albert Y. Kim"
date: "Fri Nov 11, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(mosaic)
```




## Recall LC Code for Lady Tasting Tea

```{r, eval=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)

# Simulate many, many, many times someone guessing at random
guess_cup <- c(1, 0)
simulation <- do(10000) * resample(guess_cup, size=8)

# Count number correct
simulation <- simulation %>% 
  mutate(n_correct = V1 + V2 + V3 + V4 + V5 + V6 + V7 + V8) 

# Plot
ggplot(simulation, aes(x=n_correct)) + 
  geom_bar() +
  labs(x="Number of Guesses Correct") +
  geom_vline(xintercept=8, col="red", size=1) 
```





## Recall: Definitions

> * We have two competing **hypotheses**:
>     + $H_0$: the taster is guessing at random
>     + $H_A$: the taster is not. i.e. she can really tell which came first
> * The **test statistic** used here is: # of correct guesses
> * The **observed test statistic** is 8 correct (out of 8)
> * We going to **assume** $H_0$ is true and see how likely the observed test statistic was.





## If She's Guessing at Random...

**Not very likely!**

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)

# Simulate many, many, many times someone guessing at random
guess_cup <- c(1, 0)
simulation <- do(10000) * resample(guess_cup, size=8)

# Count number correct and plot
simulation <- simulation %>% 
  mutate(n_correct = V1 + V2 + V3 + V4 + V5 + V6 + V7 + V8) 
ggplot(simulation, aes(x=n_correct)) + 
  geom_bar() +
  labs(x="Number of Guesses Correct") +
  geom_vline(xintercept=8, col="red", size=1) 
```




## Recall: Definitions

> * The bar chart of the 10000 simulations of 8 random guesses is the **null distribution**.
> * Formally: it is the distribution that describes the random behavior of the test statistic
assuming $H_0$ is true.




## Today's Single Definition

**p-value**

> * Statistically: the probability of observing a test statistic as extreme as the
one observed **assuming $H_0$ is true**.
> * Conceptually: measures our degree of surprise (lower = more surprised) of what we observed **assuming $H_0$ is true**.
> * Ex: how unlikely is it that the taster got 8 out of 8 right **assuming she was just guessing at random**




## Convert Counts to Proportions

This combines the two most difficult data manipulation/visulalization concepts in this class:

* counting the different values of `n_correct` using `n()`
* dividing by the number of simulations `10000`
* using `geom_bar()`, but now we have an explicity `y` aesthetic `prop_correct`, so we set `stat="identity"`




## Code

Please read later:

```{r, eval=FALSE}
simulation <- simulation %>% 
  group_by(n_correct) %>%
  summarise(prop_correct = n()/10000)

ggplot(simulation, aes(x=n_correct, y=prop_correct)) + 
  geom_bar(stat="identity") +
  labs(x="Number of Guesses Correct", y="Proportion") +
  geom_vline(xintercept=8, col="red", size=2) 
```




## How unlikely?

The proportion of times 8 occurs is 0.0050 = 0.5%.

```{r, echo=FALSE}
simulation <- simulation %>% 
  group_by(n_correct) %>%
  summarise(prop_correct = n()/10000)

ggplot(simulation, aes(x=n_correct, y=prop_correct)) + 
  geom_bar(stat="identity") +
  labs(x="Number of Guesses Correct", y="Proportion") +
  geom_vline(xintercept=8, col="red", size=2) 
```




## p-Value and Conclusion

p-Value meaning:

> * Conceptually: **Assuming she was guessing at random**, we are **very surprised** at observing 8 correct guesses
> * Statistically: probability of observing a test statistic of 8 **assuming $H_0$ is true** is approx. 0.5%.  i.e. TINY


Study Conclusion:

> * Conceptually: She was probably NOT guessing at random; she probably could tell.
> * Statistically: We reject $H_0$ in favor of $H_A$









