---
title: "5NG Learning Checks"
author: "Albert Y. Kim"
date: "September 21, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(lubridate)
library(knitr)

library(nycflights13)
library(ggplot2movies)
library(okcupiddata)
library(babynames)
library(fueleconomy)
library(stringr)
```



## Scatterplot

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
data(movies)
movies <- movies %>% 
  select(title, budget, rating)  

ggplot(data=movies, aes(x=budget, y=rating)) +
  geom_point() +
  labs(x="Movie Budget", y="IMDB Rating", title="IMDB Rating vs Budget")
```

### Solution

Let's look at a random sample of 5 of the movies:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
movies %>% 
  filter(!is.na(budget)) %>% 
  sample_n(5) %>% 
  kable()
```

**Both variables are numerical**. Here are the components of the Grammar of Graphics:

`data` variable | `aes()`thetic attribute  | `geom_`etric object
------------- | ------------- | -------------
`budget` | `x` | `point` 
`rating` | `y` | `point` 

### Question

Does spending more on a movie yield higher IMDB ratings?










## Linegraph

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Line graph
data(flights)
flights <- flights %>% 
  mutate(date=ymd(paste(year, month, day, sep="-"))) %>% 
  filter(month==1) %>% 
  group_by(date) %>% 
  tally()
  
ggplot(data=flights, aes(x=date, y=n)) + 
  labs(x="Date", y="Number of Flights", title="Number of Daily Flights from NYC in Jan 2013") +
  geom_line()
```

### Solution

Let's look at a random sample of 5 of the dates:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
flights %>% 
  sample_n(5) %>% 
  kable()
```

**Both variables are numerical (dates are technically numerical)**. Here are the components of the Grammar of Graphics:

`data` variable | `aes()`thetic attribute  | `geom_`etric object
------------- | ------------- | -------------
`date` | `x` | `line` 
`n` | `y` | `line` 

**Note**: Why did we use `line` as the `geom_`etric object? Because lines suggest sequence/relationship, and points don't.

### Question

Why are there drops in the number of flights?










## Boxplot

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
data(vehicles)

# Boxplot
vehicles <- vehicles %>% 
  mutate(
    trans2 = ifelse(str_sub(trans, 1, 1) == "M", "Manual", "Automatic"),
    name = paste(year, make, model)
    ) %>% 
  select(name, trans2, hwy) %>% 
  rename(trans=trans2)

ggplot(data=vehicles, aes(x=trans, y=hwy)) +
  geom_boxplot() +
  labs(y="Highway Mileage (miles/gallon)", x="Transmission Type", title="Highway Mileage of all Car Models 1984-2015") +
  coord_cartesian(ylim=c(5, 45))
```

### Solution

Let's look at a random sample of 5 of the car year/make/model matchings:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
vehicles %>% 
  sample_n(5) %>% 
  kable()
```

**`trans` type is categorical, whereas `hwy` is numerical**. Here are the components of the Grammar of Graphics:

`data` variable | `aes()`thetic attribute  | `geom_`etric object
------------- | ------------- | -------------
`trans` | `x` | `boxplot` 
`hwy` | `y` | `boxplot` 



### Question

About what proportion of manual car models sold between 1984 and 2015 got 20 mpg or worse mileage?







## Bar Plot

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
data(babynames)

# Bar Chart
babynames <- babynames %>% 
  mutate(decade = round(year / 10) * 10) %>% 
  filter(name=="Hayden" | name == "Carlos" | name == "Ethan") %>% 
  filter(decade >= 1990) %>% 
  group_by(name) %>% 
  summarise(n=sum(n))
  
ggplot(data=babynames, aes(x=name, y=n)) +
  geom_bar(stat="identity") +
  labs(x="Name", y="Count", title="Name Counts since 1990")
```

### Solution

Let's look at all the data:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
babynames %>% 
  kable()
```

**Name is categorical**. Here are the components of the Grammar of Graphics:

`data` variable | `aes()`thetic attribute  | `geom_`etric object
------------- | ------------- | -------------
`name` | `x` | `bar` 
`n` | `y` | `bar` 

### Question

About how many babies were named "Hayden" between 1990-2014?










## Histogram

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
data(profiles)

# Histogram
profiles <- profiles %>% 
  select(sex, height)

ggplot(data=profiles, aes(x=height)) +
  geom_histogram() +
  labs(x="height (in inches)", y="count", title="Heights of San Francisco OkCupid Users")
```

### Solution

Let's look at a random sample of 5 of the users:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Scatterplot i.e. bivariate plot
profiles %>% 
  sample_n(5) %>% 
  kable(row.names=FALSE)
```

**Height is numerical**. Here are the components of the Grammar of Graphics:

`data` variable | `aes()`thetic attribute  | `geom_`etric object
------------- | ------------- | -------------
`height` | `x` | `histogram` 


**Note**: We'll see later there is no `y` aesthetic here, because there is
no explicit variable that maps to it, but rather it is computed internally.

### Question

What are the smallest and largest visible heights and what do you think of them?
Also, think of one graph improvement to better convey information about SF
OkCupid users.

