---
title: "Barplots Learning Checks"
author: "Albert Y. Kim"
date: "Mon Oct 11, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(knitr)
library(nycflights13)
```



## Load Packages and Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load necessary packages and data
library(ggplot2)
library(dplyr)
library(nycflights13)

data(flights)
data(airports)
data(airlines)
```



## LC 4.27, 4.29, and 4.30

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data = flights, aes(x = carrier)) +
  geom_bar()
flights_table <- count(x = flights, vars = carrier) 
# Raw table:
flights_table
# Table sorted in descending order:
flights_table %>% 
  arrange(desc(n))
```

* Why are histograms inappropriate for visualizing categorical variables?
* How many Envoy Air flights departed NYC in 2013?
* What was the seventh highest airline in terms of departed flights from NYC in 2013?

### Solution

* Histograms are for continuous variables i.e. the horizontal part of each
histogram bar represents an interval, whereas for a categorical variable each
bar represents only one level of the categorical variable.
* Envoy Air is carrier code `MQ` and thus 26397 flights departed NYC in 2013. The
`flights_table` and `airlines` datasets should be `join`ed so that we know what these
airport codes mean!
* `US` i.e. US airways was 7th with 20536 flights. The `arrange(desc(n))` command came in real handy here!




## LC 4.31-4.32

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
flights_namedports <- inner_join(flights, airports, by = c("origin" = "faa"))
ggplot(data = flights_namedports, aes(x = carrier, fill = name)) +
  geom_bar()
```

* What kinds of questions are not easily answered by looking at the above figure?
* What can you say, if anything, about the relationship between airline and airport in NYC in 2013 in regards to the number of departing flights?

### Solution

* Because the red, green, and blue bars don't all start at 0 (only red does), it makes comparing counts hard.
* The different airlines prefer different airports. For example, United is mostly a Newark carrier and JetBlue is a JFK carrier. If airlines didn't prefer airports, each color would be roughly one third of each bar.




## LC 4.35-4.36

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
ggplot(data = flights_namedports, aes(x = carrier, fill = name)) +
  geom_bar() +
  facet_wrap(~name, ncol=1)
```

* Why is the faceted barplot preferred to the stacked barplot in this case?
* What information about the different carriers at different airports is more easily seen in the faceted barplot?

### Solution

* We can easily compare the different aiports for a given carrier using a single vertical comparison line i.e. things are lined up
* Now we can also compare the different carriers **within** a particular airport easily too. For example, we can read off who the top carrier
for each airport is easily using a single horizontal line.





