---
title: "Boxplots Learning Checks"
author: "Albert Y. Kim"
date: "Mon Oct 3, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(knitr)
library(nycflights13)
```



## Load Packages and Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load necessary packages
library(ggplot2)
library(dplyr)
library(nycflights13)

# Load weather data set in nycflights
data(weather)
```



## LC 4.17

What kind of variable is on the x-axis in this plot:

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data = weather, mapping = aes(x = factor(month), y = temp)) +
  geom_boxplot()
```

### Solution

* `month` by itself was originally coded as a **numerical variable** whose
values were `1` through `12`
* But by using `factor(month)` we are converting it to a **categorical
variable** whose "labels" are 1 through 12
* Note: we could've changed the `month` variable to a categorical one in the
data set `weather` itself, but we don't have the data manipulation tools to do
that yet!




## LC 4.18

What does the dot at the bottom of the plot for May correspond to? Explain what might have occurred in May to produce this point.

### Solution

It appears to be an outlier. Let's revisit the use of the `filter` command to hone in on it. We want all data points
where the `month` is 5 and `temp<25`

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE, fig.width=8, fig.height=4.5}
filter(weather, month==5 & temp < 25)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
filter(weather, month==5 & temp < 25) %>% kable()
```

There appears to be only one hour and only at JFK that recorded 13.1 F (-10.5 C) in the month of May. This is probably a data entry mistake!
Why wasn't the weather at least similar at EWR (Newark) and LGA (La Guardia)?



## LC 4.19

Which months tend to have the highest temperature? What reasons do you think this is?

### Solution

The solid black lines inside each box indicate **medians** (not the means!!!),
which is a measure of center. For example, over half of August data points are
about 73 F or higher. I hate heat and humidity! Blech!

As is fairly obvious, the summer months of June, July, and August are the
hottest months.




## LC 4.20

Which months tend to have the highest **variability** in temperature? What
reasons do you think this is?

### Solution

We are now interested in the **spread** of the data. One measure some of you may
have seen previously is the standard deviation. But in this plot we can read off
the Interquartile Range (IQR):

* The distance from the 1st to the 3rd quartiles i.e. the length of the boxes
* You can also think of this as the spread of the **middle 50%** of the data

Just from eyeballing it, it seems

* November has the biggest IQR, i.e. the widest box, so has the most variation in temperature
* August has the smallest IQR, i.e. the narrowest box, so is the most consistent temperature-wise

Here's how we compute the exact IQR values for each month (we'll see this more in depth Chapter 5 of the text):

1. `group` the observations by `month` then
1. for each `group`, i.e. `month`, `summarise` it by applying the summary statistic function `IQR()`, while making sure to skip over missing data via `na.rm=TRUE` then
1. `arrange` the table in `desc`ending order of `IQR`

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE, fig.width=8, fig.height=4.5}
group_by(weather, month) %>% 
  summarise(IQR = IQR(temp, na.rm=TRUE)) %>% 
  arrange(desc(IQR))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.width=8, fig.height=4.5}
group_by(weather, month) %>% 
  summarise(IQR = IQR(temp, na.rm=TRUE)) %>% 
  arrange(desc(IQR)) %>% 
  kable()
```




## LC 4.21

Create a similar boxplot as above but for `wind_speed`. What do you observe and
does this make sense?

### Solution

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data = weather, mapping = aes(x = factor(month), y = wind_speed)) +
  geom_boxplot()
```

There is clearly an outlier in February. What are the units of wind speed?  Look
at the help file of the `weather` data set by typing `?weather` in your console,
you'll see that they are in mph.

But does a wind speed of 1000 mph even make sense? Google the following: "fastest wind speed
ever recorded". Mount Washington in New Hampshire had the record of 231 mph
(recorded in 1934) up until 1996, when it was eclipsed by Barrow Island in
Australia in 1996 with a speed of 253 miles per hour (see
[here](https://www.mountwashington.org/about-us/history/world-record-wind.aspx))
during Typhoon Olivia. Not even close to 1000 mph

This is clearly a data entry mistake. What can we do to deal with this? We can either:  

**Change the range on the y-axis** using the `ylim()` command

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data = weather, mapping = aes(x = factor(month), y = wind_speed)) +
  geom_boxplot() +
  ylim(0, 45)
```

or

**Delete that observation entirely** using, again, the `filter()` command. We create a new
cleaned weather data set called `weather_cleaned` which keeps all rows with `wind_speed` less than 500 mph:

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
weather_cleaned <- filter(weather, wind_speed < 500)
ggplot(data = weather_cleaned, mapping = aes(x = factor(month), y = wind_speed)) +
  geom_boxplot()
```

