---
title: "Boxplots Learning Checks"
author: "Albert Y. Kim"
date: "Fri Sep 30, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(lubridate)
library(knitr)
library(babynames)
```



## LC 4.11

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load Packages
library(dplyr)
library(ggplot2)
library(babynames)

# Load babynames data set
data(babynames)

# Get all babies named Audrey or Jerry
all_audreys_and_jerrys <- filter(babynames, name=="Audrey" | name=="Jerry")

# Plot using facets
ggplot(data=all_audreys_and_jerrys, aes(x=year, y=prop, col=sex)) + 
  geom_line() + 
  facet_wrap(~name)
```

Flip the `facet_` and `color` variables. i.e. show a similar plot with colors representing the name and faceted by sex. Which do you think is preferable?


### Solution

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data=all_audreys_and_jerrys, aes(x=year, y=prop, col=name)) + 
  geom_line() + 
  facet_wrap(~sex)
```

Neither plot is "right" or "wrong" in any absolute sense of the word: it depends
on what comparison you want to emphasize. The only comparison I find interesting
is that while Audrey is almost exclusively female, Jerry was not exclusively
male: there was a brief period in the 1940's where some females were named
Jerry. IMO the earlier plot emphasizes this better.
