---
title: "`select()` and `filter()` Learning Checks"
author: "Albert Y. Kim"
date: "Wed Oct 12, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(knitr)
library(nycflights13)
```



## Load Packages and Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load necessary packages
library(ggplot2)
library(dplyr)
library(nycflights13)

# Load necessary data sets from nycflights
data(flights)
```



## LC 5.1, 5.3

* How many different ways are there to select all three of `dest`, `air_time`, and `distance` variables from `flights`?  Give the code showing how to do all of them you can think of.
* Why might we want to use the `select()` function on a data frame?



### Solution

We recall the variables we have by using the `names()` function:

```{r}
names(flights)
```


* We could either
    + `select(flights, dest, air_time, distance)` i.e. select them directly
    + `select(flights, dest:distance)` i.e. select a range of them, since they are sequential columns
    + `select(flights, -year, -month, -day, ETC)` i.e. deselect all the other ones. Admitedly this is rather awkward.
    + many more...
* I would probably do the first one, b/c it doesn't assume that `dest`, `air_time`, and `distance` are sequentially columns.
* `select()` would help pare down the number of columns so that we can easily `View()` them.



## LC 5.4

```{r}
not_summer_flights <- filter(flights, !between(month, 6, 8))
```

Instead of using the `!` function, what is the other way we could filter only the rows that are not summer months (June, July, or August) in the `flights` data frame? Test it out.

### Solution

Lots of different ways! Try these out!

```{r, eval=FALSE}
filter(flights, 
       month == 1 | month == 2 | month == 3 | month == 4 | month == 5 | month == 9 | month == 10 | month == 11 | month == 12)
```

This definitely not as efficient as using the `!` operator.
