---
title: "Engineering Graphics Learning Checks"
author: "Albert Y. Kim"
date: "September 23, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Data

Run the following first in your console to create the data example:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(ggplot2)

# Create data frame
simple_ex <-
  data_frame(
    A = c(1, 2, 3, 4),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("a", "a", "b", "b")
  )
```


Let's view the data frame, which is in tidy format:

```{r, echo=TRUE, eval=FALSE}
View(simple_ex)
```

```{r, echo=FALSE}
simple_ex %>% knitr::kable()
```



## Recall

* A statistical graphic is a mapping of `data` variables to `aes()`thetic
attributes of `geom_`etric objects.
* A scatterplot has `point`s as the `geom_`etric object
* A linegraph has `line`s as the `geom_`etric object



## 1. Basic Scatterplot

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data=simple_ex, aes(x=A, y=B)) + 
  geom_point()
```



## 2. Scatterplot with Color

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `color` is variable `D`

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data=simple_ex, aes(x=A, y=B, color=D)) + 
  geom_point()
```



## 3. Scatterplot with Sizes

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `size` is variable `C`

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data=simple_ex, aes(x=A, y=B, size=C)) + 
  geom_point()
```



## 4. Line Graph

* the `geom_`etric objects are `line`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data=simple_ex, aes(x=A, y=B)) + 
  geom_line()
```


## 5. Line Graph with Color

* the `geom_`etric objects are `line`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `color` is variable `D`

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data=simple_ex, aes(x=A, y=B, color=D)) + 
  geom_line()
```
