---
title: "Arrange and Joins Learning Checks"
author: "Albert Y. Kim"
date: "Fri Oct 21, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(knitr)
library(nycflights13)
```



## Load Packages and Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load necessary packages
library(ggplot2)
library(dplyr)
library(nycflights13)

# Load flights data set in nycflights
data(flights)
```





## LC 5.12

Create a new data frame that shows the top 5 airports with the largest arrival
delays from NYC in 2013.

#### Solution

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
flights %>%
  group_by(dest) %>%
  summarize(largest_arrival_delay = max(arr_delay, na.rm=TRUE)) %>%
  top_n(n = 5) %>%
  arrange(desc(largest_arrival_delay))
```

1272 minutes = 21.2 hour delay for a flight to Honolulu! So on top of the long,
long flight, you arrive nearly a day late!





## LC 5.16

What happens when you try to `left_join` the `ten_freq_dests` data frame with
`airports` instead of `airports_small`?  How might one use this result to answer
further questions about the top 10 destinations?

#### Solution

We first define the necessary data frames

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
airports_small <- airports %>%
  select(faa, name)

ten_freq_dests <- flights %>%
  group_by(dest) %>%
  summarize(num_flights = n()) %>%
  top_n(n = 10) %>%
  arrange(desc(num_flights))
```

We compare the two possible joins:

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
orig_join <- ten_freq_dests %>%
  left_join(airports_small, by = c("dest" = "faa"))
new_join <- ten_freq_dests %>%
  left_join(airports, by = c("dest" = "faa"))
```

We then do a `View()` of both:

```{r, eval=FALSE, echo=TRUE}
View(orig_join)
View(new_join)
```


The latter profiles more information. For example, most of the top 10
destinations have `tz=-5`. Looking at `?airports`, we see that `tz` corresponds
to time zone. 7 of the top 10 destinations are in the Eastern time zone, with
two more being in Pacific.




## LC 5.17

What surprises you about the top 10 destinations from NYC in 2013?

#### Solution

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ten_freq_dests %>%
  left_join(airports_small, by = c("dest" = "faa"))
```

Different people will have different answers, but I'm wondering: are that many
people flying to Boston from NYC?



