---
title: "Hypothesis Testing via Lady Tasting Tea"
author: "Albert Y. Kim"
date: "Mon Nov 7, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    df_print: kable
---



## Competing Hypotheses

Recall the two competing hypotheses:

1. She truly has the ability to tell which came first: milk or tea.
1. She is just guessing.

**Key**: Let's suppose 2 is true; suppose she is guessing.


## Learning Check

1. See if using the `resample()` and `do()` commands you can
    + simulate many, many, many cases of someone guessing at random for the eight cups, then
    + count the number of correct guesses
1. How would you compare
    + the **observed** number of correct guesses (in this case 8)
    + to the **typical** number of correct guesses assuming she is guessing at random?


## Solution

#### 1. Simulating someone guessing at random for 8 cups many, many, many times

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)

# Single cup
guess_cup <- c(1, 0)

# Simulate many, many, many times someone guessing at random
simulation <- do(10000) * resample(guess_cup, size=8)
```
```{r, echo=FALSE}
simulation %>% head()
```

#### 2. Count the number of correct guesses

```{r}
simulation <- simulation %>% 
  mutate(n_correct = V1 + V2 + V3 + V4 + V5 + V6 + V7 + V8) 
```
```{r, echo=FALSE}
simulation %>% head()
```

#### 3. What is typical number of correct guesses assuming she is guessing at random?

i.e. What is the distribution of `n_correct`?

```{r, fig.width=8, fig.height=4.5}
ggplot(simulation, aes(x=n_correct)) + 
  geom_bar() +
  labs(x="Number of Guesses Correct")
```


#### 4. Compare to observed number correct

i.e. she got 8/8 correct. We add a vertical red line:

```{r, fig.width=8, fig.height=4.5}
ggplot(simulation, aes(x=n_correct)) + 
  geom_bar() +
  labs(x="Number of Guesses Correct") +
  geom_vline(xintercept=8, col="red", size=2) 
```


<!--
#### 4. Convert to proportions 

Convert counts on y-axis to proportions by

* counting the different values of `n_correct` using `n()`
* dividing by the number of simulations `10000`
* using `geom_bar()`, but now we have an explicity `y` aesthetic `prop_correct`, so we set `stat="identity"`

```{r, fig.width=8, fig.height=4.5}
simulation <- simulation %>% 
  group_by(n_correct) %>%
  summarise(prop_correct = n()/10000)

ggplot(simulation, aes(x=n_correct, y=prop_correct)) + 
  geom_bar(stat="identity") +
  labs(x="Number of Guesses Correct", y="Proportion")
```

#### 5. Compare to observed number correct

i.e. she got 8/8 correct. We add a vertical red line:

```{r, fig.width=8, fig.height=4.5}
ggplot(simulation, aes(x=n_correct, y=prop_correct)) + 
  geom_bar(stat="identity") +
  labs(x="Number of Guesses Correct", y="Proportion") +
  geom_vline(xintercept=8, col="red", size=2) 
```
-->

## Questions

1. If she were truly guessing at random, how likely is it that she gets 8/8 correct?
1. What does the observed data say about the hypothesis that she was guessing at random?

