---
title: "Linegraphs Learning Checks"
author: "Albert Y. Kim"
date: "Wed Sep 28, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
```




## Load Necessary Data and Packages

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
library(dplyr)
library(ggplot2)
library(babynames)
```



## LC 4.7-4.8

* Why should line graphs be avoided when there is not a clear ordering of the horizontal axis?
* Why are line graphs frequently used when time is the explanatory variable?

### Solution

* Because line suggest connectedness and **correlation**. More on this later.





## LC 4.9

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
female_audreys <- filter(babynames, name=="Audrey" & sex=="F")
ggplot(data=female_audreys, aes(x=year, y=prop)) + 
  geom_line() +
  geom_smooth(se=FALSE, span=0.1)
```

* Set `span=10` in the last code block above. Does this appear to be a good smoother?

### Solution

In my opinion, `span=10` is a bit too coarse:

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data=female_audreys, aes(x=year, y=prop)) + 
  geom_line() +
  geom_smooth(se=FALSE, span=10)
```

But maybe `span=0.1` is TOO refined, i.e. not enough smoothing is happening. What about
`span=1`:

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data=female_audreys, aes(x=year, y=prop)) + 
  geom_line() +
  geom_smooth(se=FALSE, span=1)
```





## LC 4.10

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
ggplot(data=female_audreys, aes(x=year, y=prop)) + 
  geom_line() +
  geom_smooth(se=FALSE)
ggplot(data=female_audreys, aes(x=year, y=prop)) + 
  geom_line() +
  geom_smooth(method="lm", se=FALSE)
```

Do you think the earlier smoother or the regression line is a better way to pick out the "signal" (i.e. the trend) from the "noise" in the previous plot? Using this evidence, what do you think is a condition for a regression line to have "valid" interpretability?

### Solution

Hard to say.

* The regression line seems to suggest that the **overall** trend seems to be up, but doesn't fit the curve very well
* The curved smoother only picks out **recent** trends, for example the dip in the 1970's
