---
title: "Piping, Grouping, and Summarize Learning Checks"
author: "Albert Y. Kim"
date: "Fri Oct 14, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(79)
library(tidyverse)
library(knitr)
```



## Load Packages and Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Load necessary packages
library(ggplot2)
library(dplyr)
library(nycflights13)

# Load weather data set in nycflights
data(weather)
```







## LC 5.5

What does the standard deviation column in the `summary_temp_by_month` data frame tell us about temperatures in New York City throughout the year?

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
summary_temp_by_month <- weather %>% 
  group_by(month) %>% 
  summarize(
          mean = mean(temp, na.rm = TRUE),
          std_dev = sd(temp, na.rm = TRUE)
          )
```

```{r, eval=FALSE}
summary_temp_by_month
```
```{r, echo=FALSE}
kable(summary_temp_by_month)
```


### Solution

The standard deviation is a quantification of **spread** and **variability**. We
see that the period in November, December, and January has the most variation in
weather, so you can expect very different temperatures on different days.

**Note**: both `mean(temp, na.rm = TRUE)` and `sd(temp, na.rm = TRUE)` have
a `na.rm = TRUE` to ignore missing values. This should only be used when necessary 
i.e. when there actually are missing values in the data set.







## LC 5.6

What code would be required to get the mean and standard deviation temperature for each airport in NYC? Do this with and without using the `%>%` operator.

### Solution

Just switch `month` above with `origin`. First without `%>%` piping. I find this
awkward, as we first need to create an intermediate variable
`weather_group_by_airport`.

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
weather_group_by_airport <- group_by(weather, origin) 
summary_temp_by_airport <- 
  summarize(weather_group_by_airport,
          mean = mean(temp, na.rm = TRUE),
          std_dev = sd(temp, na.rm = TRUE)
          )
```


Then with `%>%` piping which IMO is less awkward. Recall `%>%` is pronounced "then".

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
summary_temp_by_airport <- weather %>% 
  group_by(origin) %>% 
  summarize(
          mean = mean(temp, na.rm = TRUE),
          std_dev = sd(temp, na.rm = TRUE)
          )
```



We output `summary_temp_by_airport`:

```{r, eval=FALSE}
summary_temp_by_airport
```
```{r, echo=FALSE}
kable(summary_temp_by_airport)
```

Is JFK significantly colder than Newark or La Guardia? Is that difference meaningful?







## LC 5.7

How could we identify how many flights left each of the three airports in each of the months of 2013?

### Solution

We could summarize the count from each airport using the `n()` function, which *counts rows*.

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
count_flights_by_airport <- weather %>% 
  group_by(origin) %>% 
  summarize(count=n())
```

```{r, eval=FALSE}
count_flights_by_airport
```
```{r, echo=FALSE}
kable(count_flights_by_airport)
```

All remarkably similar!

**Note**: the `n()` function counts rows, whereas the `sum(VARIABLE_NAME)`
funciton sums all values of a certain variable `VARIABLE_NAME`.







## LC 5.8

How could we identify the coldest temperature recorded at each airport without
using the `View()` command?

### Solution

There are three airports: `EWR`, `JFK`, and `LGA`

* Run `View(weather)`
* Click on "Filter"
* Under `origin` type in `EWR`
* The click the arrow next to `temp` to get the coldest day. For example for `EWR` it was 10.94 degrees
* Repeat for all remaining airports
