---
title: "p-Values"
author: "Albert Y. Kim"
date: "Mon Nov 14, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

## To be Completed by Wednesday 11am



## Even vs Odd # of Letters in Last Name

Say I give you a data set with two variables: 1) even or odd number of letters
in last name i.e. two groups and 2) test score. For example, here is an example
of what this data would look like, with only 3 rows:

```{r, echo=FALSE, message=FALSE}
library(dplyr)
data_frame(
  id = c(1, 2, 3),
  num_letters = c("even", "odd", "odd"),
  test_score = c(0.7, 0.6, 0.8)
)
```

Think about how you would test:

* $H_0$: no difference in test scores between odd vs even
* $H_A$: there is a difference

Hints:

* Think about what assuming $H_0$ allows you to do the data set
* Question from Problem Set 08 Question 1.b)



#### Assuming $H_0$ Allows You To...

All hypothesis is done **assuming the null hypothesis is true**. In our case, we assume
there is no difference in test scores between those with even # and those with odd #.

So for the above data, for each student, it doesn't matter if they have an even or odd #. In other
words, the variable `num_letters` is meaningless. If it's meaningless, then we can permute it
and nothing should matter!

Thus assuming $H_0$ is true, the above data is the same as the following data

```{r, echo=FALSE, message=FALSE}
data_frame(
  id = c(1, 2, 3),
  num_letters = c("odd", "even", "odd"),
  test_score = c(0.7, 0.6, 0.8)
)
```

which is the same as the following data

```{r, echo=FALSE, message=FALSE}
data_frame(
  id = c(1, 2, 3),
  num_letters = c("odd", "odd", "even"),
  test_score = c(0.7, 0.6, 0.8)
)
```



#### Test Statistic

Now we need a **test statistic**. A statistic is just a summary of many values
of numbers to a single one. A test statistic is simply a statistic used for the
purpose of testing.

Here, we want to compare odd vs even. So what's an obvious test statistic?
The mean test score of odd MINUS the mean test score of even. Or using mathematical notation:
$\mu_E- \mu_O$. $\mu$ is the traditional character for means.

What is the **observed test statistic**? $\overline{x}_E - \overline{x}_o$: the
observed difference in means.



#### Null Distribution

The null distribution is the distribution of the test statistic assuming $H_0$ is
true. How do we construct this? Using `shuffle()`!














