---
title: "Probability via Simulation"
author: "Albert Y. Kim"
date: "Mon Oct 31, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(mosaic)
```


* LC1: Create a histogram of the number illustrating the long-run behavior of flipping a coin 10 times. 
    + Where is it centered?
    + Describe the shape of the **distribution** of values
* LC2: Tie all elements of Line 13 of code below to the Attributes of Powerball seen in the slides above.
* LC3: Try to replicate the above, but for the sum of two die rolls. Hint: `resample(c(1:6), 2)`



## LC1

```{r}
coin_flips <- do(500)*rflip(10)
coin_flips <- coin_flips %>% 
  as_data_frame()

ggplot(coin_flips, aes(x=heads)) +
  geom_histogram(binwidth = 1)
```



* Where is it centered?
* Describe the shape of the **distribution** of values



## LC2

* Atrributes of the Lottery Machine:
    * How many balls do you have? **2**
    * What are written on the balls? **Heads and tails**
    * Do the balls have equal probability of being picked? **Yes**
* Attributes of the Drawing:
    * How are you drawing the balls? **At random**
    * How many balls do you draw? **10**
    * What are you recording about each drawn ball? **If heads or tails**
    * What do you do with drawn balls? **You DON'T put them back**
* Number of Lotteries:
    * How many times do you repeat the lottery? **Many many many times. In the above case, 500 times**



## LC3

```{r}
two_dice <- do(500) * sample(c(1:6), 2)
two_dice <- two_dice %>% 
  as_data_frame() %>% 
  mutate(sum = V1 + V2)

ggplot(two_dice, aes(x=sum)) +
  geom_bar()
```