---
title: "Probability via Simulation"
author: "Albert Y. Kim"
date: "Mon Oct 31, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(mosaic)
```


* LC1: Create a histogram of the number illustrating the long-run behavior of flipping a coin 10 times. 
    + Where is it centered?
    + Describe the shape of the **distribution** of values
* LC2: Tie all elements of Line 13 of code below to the Attributes of Powerball seen in the slides above.
* LC3: Try to replicate the above, but for the sum of two die rolls. Hint: `resample(c(1:6), 2)`



## LC1

```{r}
coin_flips <- do(500)*rflip(10)
coin_flips <- coin_flips %>% 
  as_data_frame()
```

If we `View(coin_flips)` the first 6 rows, we see that we have in **tidy format**:

```{r, echo=FALSE}
coin_flips %>% 
  head() %>% 
  knitr::kable()
```

So we plot a histogram of the `heads` variable with `binwidth=1` since we are
dealing with integers i.e. whole numbers.

```{r}
ggplot(coin_flips, aes(x=heads)) +
  geom_histogram(binwidth = 1)
```

* Where is it centered? Answer: At 5 i.e. half of 10.
* Describe the shape of the **distribution** of values. Answer: bell-shaped.
i.e. like a Normal distribution.






## LC2

* Atrributes of the Lottery Machine:
    * How many balls do you have? **2**
    * What are written on the balls? **Heads and tails**
    * Do the balls have equal probability of being picked? **Yes**
* Attributes of the Drawing:
    * How are you drawing the balls? **At random**
    * How many balls do you draw? **10**
    * What are you recording about each drawn ball? **If the ball has heads or tails**
    * What do you do with drawn balls? **You put them back in the machine**
* Number of Lotteries:
    * How many times do you repeat the lottery? **Many many many times. In the above case, 500 times**



## LC3

Let's unpack `sample(c(1:6), 2)`:

* Running `c(1:6)` in the console returns six values, `1 2 3 4 5 6`, one for each possible die roll value.
* `sample(c(1:6), 2)` says: sample a value from 1 to 6 twice. This is akin to rolling a die twice.

```{r}
two_dice <- do(500) * sample(c(1:6), 2)
two_dice <- two_dice %>% 
  as_data_frame() 
```

If we `View(two_dice)` the first 6 rows, we see that we have in **tidy format**:

```{r, echo=FALSE}
two_dice %>% 
  head() %>% 
  knitr::kable()
```

So to get the sum of the two dice, we `mutate()` a new variable `sum` based on the sum of the two die:

```{r}
two_dice <- two_dice %>% 
  mutate(sum = V1 + V2)
```

```{r, echo=FALSE}
two_dice %>% 
  head() %>% 
  knitr::kable()
```

And now we plot it:

```{r, echo=FALSE, message=FALSE}
ggplot(two_dice, aes(x=sum)) +
  geom_histogram(binwidth=1)
```


* Atrributes of the Lottery Machine:
    * How many balls do you have? **6**
    * What are written on the balls? **Integers 1 through 6**
    * Do the balls have equal probability of being picked? **Yes**
* Attributes of the Drawing:
    * How are you drawing the balls? **At random**
    * How many balls do you draw? **2**
    * What are you recording about each drawn ball? **The sum of the balls**
    * What do you do with drawn balls? **You put them back in the machine**
* Number of Lotteries:
    * How many times do you repeat the lottery? **Many many many times. In the above case, 500 times**


#### Advanced

What's the deal with the ugly axes tick marks? This is again b/c computers are stupid,
and ggplot does not know we are dealing only with whole numbers i.e. integers. We could cheat it
and treat the sum, a numerical variable, as a categorical variable using `geom_bar()`

```{r, echo=FALSE, message=FALSE}
ggplot(two_dice, aes(x=sum)) +
  geom_bar()
```

Here we can at least separate out the individual whole numbers, but we still
have the axes tick marks problem? How do I fix these? That's for the more
advanced data science class [MATH 216 Data
Science](https://rudeboybert.github.io/MATH216/).
