---
title: "Different Types of Sampling"
author: "Albert Y. Kim"
date: "Wed Nov 2, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(mosaic)
```



#### LC1

Rewrite `rflip(10)` using the `resample()` command. Hint: `coin <- c("H", "T")`

**Answer**: `resample()` by default samples **with** replacement and with equal probability. All we need to change
is the `size` argument:

```{r, eval=FALSE}
coin <- c("H", "T")

# Compare the following multiple times:
resample(coin, size=10)
rflip(10)
```


#### LC2

Rewrite the `shuffle()` command by changing the minimal number of
default settings of `resample()`. Test this on `fruit`

**Answer**: `shuffle()` is the same as sampling **without** replacement. All you need to
change the `replace` argument from the default of `TRUE` to `FALSE`.

```{r, eval=FALSE}
fruit <- c("apple", "orange", "mango")

# Compare the following multiple times:
resample(fruit, replace=FALSE)
shuffle(fruit)
```


#### LC3

Write code that will allow you to generate a sample of 15 fruit
without replacement.

**Answer**: The following yields an error. You can't sample more balls **without**
replacement than there are balls in the machine. Alternatively, the largest
sample **without** replacement of `fruit` is of size 3.

```{r, eval=FALSE}
fruit <- c("apple", "orange", "mango")
resample(fruit, size=15, replace=FALSE)
```



#### LC4

Write code that will allow you to generate a sample of 15 fruit
with replacement.

**Answer**: This works!

```{r, eval=FALSE}
fruit <- c("apple", "orange", "mango")
resample(fruit, size=15)
```


#### LC5

What's the fastest way to do the above 5 times? Write it out

**Answer**: Use `do()`!

```{r, eval=TRUE}
fruit <- c("apple", "orange", "mango")
do(5) * resample(fruit, size=15)
```
