---
title: "Residuals"
author: "Albert Y. Kim"
date: "Wed Dec 7, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

## Alaska Flights

#### Load Alaska Flights data:

```{r, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(nycflights13)
library(knitr)
data(flights)

# Load Alaska data, deleting rows that have missing dep or arr data
alaska_flights <- flights %>% 
  filter(carrier == "AS") %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay))

# Instead of looking at all 709 Alaska flights, let's look at a random sample
# of 10 of them:
set.seed(76)
alaska_flights_sample <- alaska_flights %>% 
  sample_n(10)

# Plot the 25 pairs of points and regression line:
ggplot(data=alaska_flights_sample, aes(x = dep_delay, y = arr_delay)) + 
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```



## Regression:

Let's take

* The **outcome** variable y to be departure delay `dep_delay`
* The **predictor** variable x to be arrival delay `arr_delay`


#### Running Regression

The way you run regression in R is via the `lm()` command for linear model and
assign this to `model`

```{r, eval=FALSE}
lm(OUTCOME_VARIABLE ~ PREDICTOR_VARIABLE, data=DATA_SET_NAME)
```

Let's run the regression using only the 10 randomly sampled flights:

```{r, warning=FALSE, echo=TRUE, eval=TRUE}
model <- lm(arr_delay ~ dep_delay, data=alaska_flights_sample)
```


#### Looking at Output

There are two useful functions from the broom package to look at the output of a regression:

* `tidy()` to get the **regression table**
* `augment()` to get **point-by-point** values. This table also contains more advanced output

```{r, warning=FALSE, echo=TRUE, eval=TRUE, fig.width=16/2, fig.height=9/2}
library(broom)
regression_table <- tidy(model, conf.int = TRUE)
kable(regression_table, digits=3)

point_by_point <- augment(model) %>% 
  as_data_frame()
kable(point_by_point, digits=3)
```



## Learning Check

1. Identify in the point-by-point table which columns correspond to
    + The observed outcome variable $y$: `arr_delay`
    + The fitted values on the regression line $\widehat{y}$: `.fitted`
    + The residuals and which two columns they are computed based on: `.resid`
1. Re-run the regression, but using **all** 907 flights in the `alaska_flights` 
data set and plot a histogram of the residuals. What shape does the distribution
have?


#### Residual Plots

Before we plot the histogram of residuals, let's plot the raw residuals. Imagine
we take the red line in the left-hand plot and flatten it out in the right hand
plot. This is equivalent to computing the residuals.

We see that the residuals just form random scatter/noise around the $y=0$ line i.e.
**there is no systematic pattern in the residuals**

```{r, echo=FALSE,  fig.width=16/2, fig.height=9/2}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

p1 <- ggplot(data=point_by_point, aes(x = dep_delay, y = arr_delay)) + 
  geom_point() +
  geom_smooth(method="lm", se=FALSE, col="blue") +
  labs(x="Departure Delay", y="Arrival Delay", title="Arrival Delay over Departure Delay")
p2 <- ggplot(data=point_by_point, aes(x = dep_delay, y =.resid )) + 
  geom_point() +
  geom_hline(yintercept=0, col="blue", size=1) +
  labs(x="Departure Delay", y="Residuals", title="Residuals over Departure Delay")
multiplot(p1, p2, cols=2)
```



The residuals are $\epsilon_i = \mbox{observed} - \mbox{fitted} = y_i -
\widehat{y}_i$. They roughly follow a Normal distribution centered at 0. i.e.
on average the error is 0. **This is good!**

```{r, warning=FALSE, echo=TRUE, eval=TRUE, fig.width=16/2, fig.height=9/2}
model <- lm(arr_delay ~ dep_delay, data=alaska_flights)
point_by_point <- augment(model) %>% 
  as_data_frame()
ggplot(point_by_point, aes(x=.resid)) +
  geom_histogram(binwidth = 10)
```

