---
title: "Back to Confidence Intervals"
author: "Albert Y. Kim"
date: "Thu Dec 2, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

Load the familiar data again, removing individuals with no listed height

```{r, echo=TRUE, message=FALSE}
library(mosaic)
library(dplyr)
library(ggplot2)
library(okcupiddata)
data(profiles)
set.seed(76)
profiles <- profiles %>% 
  filter(!is.na(height))
```

We take many, many, many samples of size 5 and then take the sample mean, and then do the same for samples of size 50:

```{r, echo=TRUE, message=FALSE, eval=TRUE, cache=TRUE}
n <- 5
samples_5 <- do(10000) * 
  mean(resample(profiles$height, size=n, replace=TRUE))
samples_5 <- samples_5 %>% 
  as_data_frame() 

n <- 50
samples_50 <- do(10000) * 
  mean(resample(profiles$height, size=n, replace=TRUE))
samples_50 <- samples_50 %>% 
  as_data_frame() 
```


## Learning Checks

Construct 95% confidence intervals for $\mu$, the true average height of all 60K OkCupid users

* Using both methods for constructing confidence intervals: quantiles and bell-curve aproach
* For both when `n=5` and `n=50`

Hint: Look at the histograms of the 10000 simulations.



## Notes

#### Quantiles

To get 95% confidence interval, we need

* the middle 95% (blue)
* so the tails represent 5% (red), i.e. 2.5% each
* thus we need the 2.5% and 97.5% percentiles
* thus we use `quantile()` with `prob=c(0.025, 0.975)`

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, fig.height=9/2, fig.width=16/2}
dnorm_limit <- function(x) {
    y <- dnorm(x, mean = 5, sd = 2)
    y[x <= 1  |  x >= 9] <- NA
    return(y)
}
dnorm_limit_left <- function(x) {
    y <- dnorm(x, mean = 5, sd = 2)
    y[x >= 1  ] <- NA
    return(y)
}
dnorm_limit_right <- function(x) {
    y <- dnorm(x, mean = 5, sd = 2)
    y[x < 9  ] <- NA
    return(y)
}

# ggplot() with dummy data
ggplot(data.frame(x=c(-5, 13)), aes(x=x)) + 
  stat_function(fun=dnorm_limit, geom="area", fill="blue", alpha=0.2, n=10000) +
  stat_function(fun=dnorm_limit_left, geom="area", fill="red", alpha=0.2, n=10000) +
  stat_function(fun=dnorm_limit_right, geom="area", fill="red", alpha=0.2, n=10000) +
  stat_function(fun=dnorm, args = list(mean = 5, sd = 2))
```

#### Normal Curve

We need the mean and the standard deviation of the **sampling distribution**,
i.e. the distribution of the 10,000 $\overline{x}$. The latter is the **standard
error**.



## n=5

**Quantiles**: Using the fact above

```{r}
quantile(samples_5$mean, prob=c(0.025, 0.975))
```

i.e. we have **(64.8, 71.8)**

**Normal-Curve Approach**: 

```{r}
samples_5 %>% 
  summarise(mean=mean(mean), SE=sd(mean))
```

So using the +/- 2 SD from the mean rule, we have: $(68.29 - 2\times 1.80, 68.29 + 2\times 1.80)$ = **(64.69, 71.89)**.

We plot the resulting confidence intervals i.e. **our net**

* Blue: Quantile contstructed
* Red: Normal curve constructed

```{r, message=FALSE, warning=FALSE, eval=TRUE, fig.height=9/2, fig.width=16/2}
ggplot(samples_5, aes(x=mean)) +
  geom_histogram(binwidth = 1) +
  xlim(c(60,76)) +
  geom_vline(xintercept = 68.29, linetype="dashed") +
  geom_vline(xintercept = c(64.8,  71.6), col="blue") +
  geom_vline(xintercept = c(64.69, 71.89), col="red")
```


## n=50

**Quantiles**: Using the fact above

```{r}
quantile(samples_50$mean, prob=c(0.025, 0.975))
```

i.e. we have **(67.18, 69.38)**

**Normal-Curve Approach**:

```{r}
samples_50 %>% 
  summarise(mean=mean(mean), SE=sd(mean))
```

So using the +/- 2 SD from the mean rule, we have: $(68.30 - 2\times 0.57, 68.30 + 2\times 0.57)$ = **(67.16, 69.44)**.

We plot the resulting confidence intervals i.e. **our net**

* Blue: Quantile contstructed
* Red: Normal curve constructed

```{r, message=FALSE, warning=FALSE, eval=TRUE, fig.height=9/2, fig.width=16/2}
ggplot(samples_50, aes(x=mean)) +
  geom_histogram(binwidth = 0.5) +
  xlim(c(60,76)) +
  geom_vline(xintercept = 68.30, linetype="dashed") +
  geom_vline(xintercept = c(67.18, 69.40), col="blue") +
  geom_vline(xintercept = c(67.16, 69.44), col="red")
```



## Conclusion

As expected, confidence interval is narrower for `n=50`. This is because the SE is smaller when `n=50`.






