---
title: "Standard Errors"
author: "Albert Y. Kim"
date: "Wed Nov 30, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

Let's revisit the OkCupid profile data. Run the following in your console:

```{r, echo=TRUE, message=FALSE}
library(mosaic)
library(dplyr)
library(ggplot2)

library(okcupiddata)
data(profiles)

# Added line: let's remove all users (i.e. rows) who did not list a height. is.na() returns
# TRUE if missing, so we want those that are NOT missing.
profiles <- profiles %>% 
  filter(!is.na(height))
```

Then run the following:

```{r, echo=TRUE, message=FALSE, eval=FALSE}
n <- 5
samples <- do(10000) * 
  mean(resample(profiles$height, size=n, replace=TRUE))
samples <- samples %>% 
  as_data_frame() 
ggplot(samples, aes(x=mean)) +
  geom_histogram(binwidth = 1) +
  xlim(c(40,90))
```


## Learning Checks

1. Explicity compute the **standard error** when taking samples of size `n=5` and `n=50`. 
2. Discuss with your peers **why** they matter in any study that involves some kind of sampling.


#### LC1

This code

* Samples `n=5` OkCupid users with replacement from the **population** of OkCupid users
* Computes the sample mean height $\overline{x} = \frac{1}{n}\sum_{i=1}^{n}x_i$, where $x_i$ is the height of the $i$th sampled user out of $n$.
* Does this many, many, many times i.e. 10000 times
* Plots the **distribution** of these 10000 values of $\overline{x}$ via a histogram


#### LC2

Varying `n` corresponds to sampling a different amount of people. Here are the histogram of sample means when 

* `n=5`
* `n=50`

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, cache=TRUE, fig.height=9/2, fig.width=16/2}
n <- 5
samples_5 <- do(10000) * 
  mean(resample(profiles$height, size=n, replace=TRUE))
samples_5 <- samples_5 %>% 
  as_data_frame() %>% 
  mutate(n="n=5")

n <- 50
samples_50 <- do(10000) * 
  mean(resample(profiles$height, size=n, replace=TRUE))
samples_50 <- samples_50 %>% 
  as_data_frame()%>% 
  mutate(n="n=50")

samples <- bind_rows(
  samples_5, samples_50
)

ggplot(samples, aes(x=mean)) +
  facet_wrap(~n) + 
  geom_histogram(binwidth = 1) +
  xlim(c(40,90))
```


#### LC3

The histogram for `n=500` latter is **narrower** i.e. less variable i.e. more **precise**.




## Extra

#### Point Estimate of True Population Mean

Recall that in both cases, the sample mean $\overline{x}$ is a **point
estimate** of the **true population mean $\mu$**, i.e. the true mean height of
all 60K OkCupid users. i.e. $\mu$ = `r mean(profiles$height) %>%
round(2)` inches (or about 5'8''). Let's plot this value in red:

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, cache=TRUE, fig.height=9/2, fig.width=16/2}
ggplot(samples, aes(x=mean)) +
  facet_wrap(~n) + 
  geom_histogram(binwidth = 1) +
  xlim(c(40,90)) + 
  geom_vline(xintercept = mean(profiles$height, na.rm=TRUE), col="red")
```

We see that using `n=50`, our sample means $\overline{x}$ i.e. our point 
estimates of $\mu$ are more often closer to the true population mean in red.
**We are more often closer to the true value**. This is why sample size matters!


#### Confidence Interval

Let's look at the middle 95% of values for `n=5`. It is [65.0, 71.6]. i.e.

* 2.5% of values are to the left of the left dashed line
* 95% of values are between the two dashed lines
* 2.5% of values are to the right of the right dashed line

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, cache=TRUE, fig.height=9/2, fig.width=16/2}
ggplot(samples_5, aes(x=mean)) +
  geom_histogram(binwidth = 1) +
  xlim(c(40,90)) + 
  geom_vline(xintercept = mean(profiles$height, na.rm=TRUE), col="red") +
  geom_vline(xintercept = quantile(samples_5$mean, probs=c(0.025, 0.975)), linetype="dashed")
```

Let's look at the middle 95% of values for `n=50`. It is [67.18, 69.40]:

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, cache=TRUE, fig.height=9/2, fig.width=16/2}
ggplot(samples_50, aes(x=mean)) +
  geom_histogram(binwidth = 1) +
  xlim(c(40,90)) + 
  geom_vline(xintercept = mean(profiles$height, na.rm=TRUE), col="red") +
  geom_vline(xintercept = quantile(samples_50$mean, probs=c(0.025, 0.975)), linetype="dashed")
```
