---
title: "Finishing Confidence Intervals"
author: "Albert Y. Kim"
date: "Fri Dec 3, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

Load the familiar data again, removing individuals with no listed height

```{r, echo=TRUE, message=FALSE}
library(mosaic)
library(dplyr)
library(ggplot2)
library(okcupiddata)
data(profiles)
profiles <- profiles %>% 
  filter(!is.na(height))
```



## Learning Checks

* Let's **pretend** we don't know the true mean mu = 68.3 inches
* Using the `resample()` function, take a **single sample** (without replacement) of size $n=50$ of 50 OkCupid users' heights `profiles$height`. Assign this to an object `sample_50`.
* Using the `mean()`, `sd()`, and `sqrt()` functions, compute one confidence interval for mu.
* Did your net catch the fish?



## LC1: A Single Sample

```{r}
set.seed(76)
sample_50 <- resample(profiles$height, size=50, replace = FALSE)
```



## LC2: A 95% CI

```{r}
xbar <- mean(sample_50)
s <- sd(sample_50)
n <- length(sample_50)
c(xbar, s, n)
c(xbar -2*s/sqrt(n), xbar +2*s/sqrt(n))
```

Our 95% CI is 
$$
\left(\overline{x} - 2 SE, \overline{x} + 2 SE\right) =\left(\overline{x} - 2 \frac{s}{\sqrt{n}}, \overline{x} + 2 \frac{s}{\sqrt{n}}\right) = \left(67.01, 69.67\right)
$$


## Long-Run Performance of CI

**Back to theoretical/rhetorical land** Let's repeat the following procedure 100 times:

1. Take a sample of size $n=50$
1. Compute the 95% CI based on $\overline{x}$, $s$, $\sqrt{n}$
1. See if we caught the fish


Here are the (random) results:

```{r, echo=FALSE}
n_sim <- 100
pop_mean <- mean(profiles$height)

CI <- data_frame(
  ID = 1:n_sim,
  left=0, 
  right=0,
  captured=""
)
set.seed(79)
for(i in 1:n_sim){
  sample_50 <- resample(profiles$height, size=50, replace = FALSE)
  xbar <- mean(sample_50)
  s <- sd(sample_50)
  n <- length(sample_50)
  CI[i, c("left", "right")] <- c(xbar - 2*s/sqrt(n), xbar + 2*s/sqrt(n))
  CI$captured[i] <- between(pop_mean, xbar - 2*s/sqrt(n), xbar + 2*s/sqrt(n))
}

CI <- CI %>% 
  #mutate(captured = ifelse(captured, "black", "red")) %>% 
  tidyr::gather(type, values, c(2,3))


ggplot(CI, aes(y=ID, x=values, group=ID, col=captured)) +
  geom_line(size=1) +
  geom_vline(xintercept=68.3, linetype="dashed") +
  labs(title="Real Mean Height is Dashed Line", y="", x="height")
```

Our net missed the fish 3 times! On average, it will miss it 5% of the time.


## Correct Interpretation of Original CI

So the correct interpretation of our original 95% confidence interval (67.01, 69.67)

* **IS**: the procedure that generated the CI (67.01, 69.67) is **95% reliable**.
i.e. 95% of the time it will get it right
* **IS NOT**: the probability that (67.01, 69.67) contains the true mean height is 95%.  
i.e. the probability is either 1 or 0: either is does or doesn't.


