---
title: "Residuals"
author: "Albert Y. Kim"
date: "Wed Dec 07, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(nycflights13)
data(flights)
```




## Final Note about Confidence Intervals

* **ALL** confidence intervals have form: $\mbox{Point Estimate}\pm 2 \times SE$
* What the point estimate and SE are will change


> * CI for $\mu$:
>     + Point estimate: $\overline{x}$
>     + SE = $\frac{s}{\sqrt{n}}$
>     + 95% CI: $\overline{x} \pm 2 \frac{s}{\sqrt{n}} = \left(\overline{x} - 2 \frac{s}{\sqrt{n}}, \overline{x} + 2 \frac{s}{\sqrt{n}}\right)$
> * PS-11: CI for $\mu_1 - \mu_2$ i.e. Difference in means of two groups
>     + Point estimate: $\overline{x}_1 - \overline{x}_2$
>     + SE = NASTY!


## Correlation Coefficient

* Measures the strength of linear association between two variables
* Always between [-1, 1] where
    + -1 indicates perfect negative relationship
    + 0 indicates no relationship
    + +1 indicates perfect positive relationship



## Correlation Coefficient

```{r, echo=FALSE, warning=FALSE}
library(mvtnorm) 
correlation <- c(-0.9999, -0.9, -0.75, -0.3, 0, 0.3, 0.75, 0.9, 0.9999)
n_sim <- 100

values <- NULL
for(i in 1:length(correlation)){
  rho <- correlation[i]
  sigma <- matrix(c(5, rho*sqrt(50), rho*sqrt(50), 10), 2, 2) 
  sim <- rmvnorm(
    n=n_sim,
    mean=c(20,40),
    sigma=sigma
    ) %>%
    as_data_frame() %>% 
    mutate(correlation=round(rho,2))
  
  values <- bind_rows(values, sim)
}

ggplot(data=values, aes(V1,V2)) +
  geom_point() +
  facet_wrap(~correlation, ncol=5) +
  labs(x="", y="", title="Different Correlation Coefficients") + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  )
```



## Correlation Coefficient

Two versions: Just like with $\mu$ and $\overline{x}$

* **Population correlation coefficient** $\rho$
* **Sample correlation coefficent** $r$ based on a sample of n **pairs** of observations



## Example

Recall the `nycflights` data set. For Alaska Air flights, let's explore the **relationship** between

* Departure delay
* Arrival delay

```{r, warning=FALSE, eval=FALSE}
library(nycflights13)
data(flights)

# Load Alaska data, deleting rows that have missing dep or arr data
alaska_flights <- flights %>% 
  filter(carrier == "AS") %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay))

ggplot(data=alaska_flights, aes(x = dep_delay, y = arr_delay)) + 
   geom_point()
```



## Example

```{r, warning=FALSE, echo=FALSE}
library(nycflights13)
data(flights)
alaska_flights <- flights %>% 
  filter(carrier == "AS") %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay))
ggplot(data=alaska_flights, aes(x = dep_delay, y = arr_delay)) + 
   geom_point()
```



## Example

The correlation coefficient is computed as follows:

```{r, warning=FALSE, echo=TRUE}
cor(alaska_flights$dep_delay, alaska_flights$arr_delay)
```

This is fairly strongly **positively associated**!



## Bored? 

Play [Guess the Correlation](http://guessthecorrelation.com/)



## Important Note

**Correlation coefficent $\neq$ slope of regression line.**
Example: Say we have 3 group of points

```{r, echo=FALSE, warning=FALSE}
vals <- seq(-2, 2, length=20)
example <- data_frame(
  x=rep(vals, 3),
  y=c(0.01*vals, 1*vals, 3*vals),
  slope=factor(rep(c(0.01, 1, 3), each=length(vals)))
)
ggplot(example, aes(x=x, y=y, col=slope)) +
  geom_point(size=2) +
  theme(legend.position="none")
```



## Important Note

Their regression lines have different slopes, but $r= 1$ for all 3. i.e. perfect (positive) linear relationship.

```{r, echo=FALSE, warning=FALSE}
ggplot(example, aes(x=x, y=y, col=slope)) +
  geom_point(size=2) + 
  geom_smooth(method="lm", se=FALSE)
```



## Learning Check











