---
title: "Midterm III Review"
author: "Albert Y. Kim"
date: "Mon Nov 14, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```




## Format

* Wed Nov 16, 7:30pm-10:00pm in Warner 506/507.
* Target median completion time is about 1h15m
* Closed book, no calculators, but you may bring `dplyr` cheatsheet.
* You won't need to write 100% correct R code, but rather rough **pseudocode**




## Sources

* Suggested approach
    + Read the slides from each lecture to get the **executive summary**
    + Chalk talks went slow for reason: they are important concepts not worth rushing
    + Learning check discussions
    + Problem Sets



## Topics

* Focus on Lec17 through Lec24. For Lec24 ignore anything involving p-values: Slides 6-9.
* However the midterm is still cumulative in that `ggplot2`/`dplyr` are now not
goals in and of themselves (like with Data Science), but rather the **tools** we
use for
    + Performing simulations, studying sampling, etc...
    + Examining results



## Topics

* Note the paradigm shift:
    + PS-07: recap of the tools of data visualization/manipulation **to the end of data science**
    + PS-08: using the tools of data visualization/manipulation **to the end of sampling and probability**
* Think of reading/writing as a **tool for communication** for various essay-based classes
* Much in the same way, we are using coding as the language for expressing statistical ideas.




## Extra Resources

Especially for the section 3 of the syllabus: Statistical Inference, look at the
OpenIntro textbook from Problem Set 08 and use the index to find the topics
(mostly at the front of the text).




## Some Important Topics

* Definitions relating to sampling. Ex: 
    + Tub of rice
    + LC: Comment on the generalizability/representativeness of situations
* Being able to describe any sampling situation in terms of the Powerball analogy
* Components of designing experiments. Ex: Fried chicken face off
* Defintions relating to hypothesis testing. Ex: Lady Tasting Tea




## Computational Tools

The **tools** for sampling, randomization, and probability from the `mosaic` 
package:

* `rflip()` a coin
* `shuffle()` 
* `do()` many, many, many times
* `resample()`: the swiss army knife. Understand the three knobs you can control to have different types of sampling.
