---
title: "Midterm II Review"
author: "Albert Y. Kim"
date: "Mon Oct 24, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```




## Format

* Wed Oct 23, 7:30pm-10:00pm, in Warner 506.
* I'm going to try to target it so median completion time is about 1h15m
* Closed book, no calculators, but you may bring `dplyr` cheatsheet.
* You won't need to write 100% correct R code, but rather rough **pseudocode**




## Pseudocode & Algorithms

> * [Pseudocode](http://image.slidesharecdn.com/pseudocode-111128151539-phpapp02/95/introduction-to-pseudocode-38-638.jpg?cb=1393513763)
is informal and rough code that doesn't necessarily need to work, but still
illustrates each **step of your algorithm**.
> * An **algorithm** is just a computer recipe: a process or set of rules to be followed in
calculations or other problem-solving operations.




## Sources

* Lectures 01 through 16 inclusive
    + Slides from each lecture
    + Corresponding textbook material
    + Learning check discussions
* Problem Sets




## Sources: Problem Sets

* Problem Sets 01-05: Go through them all. You are now in a position to understand all data manipulations.
* Problem Set 06 as practice for data manipulation. Instructions:
    + Separate out **what** you are going to do from **how** you are going to do it. i.e. set up a plan
    + I **highly** recommend you work in groups for this, especially the brainstorming stage.




## Data Visualization

> * For any kind of situation/data, be able to identify which of the 5NG (5 named graphs) is most 
appropriate to convey information
> * For each of the 5NG understand the **Grammar of Graphics**: `data`, `aes()`, `geom_WHATEVER`
> * How can faceting help?
> * Be able to both:
> * A: Forward engineer: I give you tidy data, you write out a rough `ggplot()` call and/or draw the graph
> * B: Reverse engineer: I give you the graph, you write out a rough `ggplot()` call and/or the tidy data




## Data Manipulation

> * Understand the 5MV (5 main verbs) + joins. The images on the `dplyr`
cheatsheet illustrate these well.
> * IMO the best way to study these to learn by doing. Go over any example of data manipulation in
the learning checks, the textbook, and Problem Sets and see if you can reconstruct them on your own.
> * If you can get them working in R, then you're **definitely** able to write the pseudocode.
