---
title: "Midterm I Review"
author: "Albert Y. Kim"
date: "Mon Oct 3, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```



## Format

* In-class Wed Oct 5. Midterms II & III will be evening midterms.
* Closed book, no calculators
* More conceptual in nature
* You won't need to write explicit R code, but you may need to understand it



## Sources

* Lectures 01 through 09 inclusive
    + Slides from each lecture
    + Corresponding textbook material
    + Learning check discussions!
* Nothing on RStudio, R Markdown, nor R packages



## Major Topics

* Tidy data. What are the components?
* What is the grammar of graphics?
* What are the Five Named Graphs (5NG)? What are their distinguishing features
* Explicitly tie the grammar of graphics to scatterplots and line graphs (no boxplots)





