
---
title: "Problem Set 07"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 4, 2016"
output:
  html_document
---



```{r, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
```




## Question 1: Women and Children First

**Research question**: Did the "women-and-children" first
policy hold? Hint: the answer is yes.

Argue your conclusion using the results from an outputed table. Do as much as
you can computationally to minimize the number of hand calculations you may need
to make (if any).

```{r, echo=FALSE, message=FALSE}
data(Titanic)
# Convert Titanic data to tidy format
Titanic <- Titanic %>% 
  as.data.frame()

Titanic %>% 
  group_by(Sex, Age, Survived) %>% 
  summarise(Freq=sum(Freq))

```

**Your conclusion**:







## Question 2: America Runs on Starbucks?

**Research question**: Do Starbucks tend to locate in higher income neighbors,
while Dunkin Donuts does not i.e. they tend to locate in lower income
neighborhoods." 

Recreate the plot in part a), but with any improvements you think are warranted 
**in order to answer the question**. Don't worry about making the green/orange
colors match.

```{r, echo=FALSE, message=FALSE}
# Copy and paste the line of code that loads DD_vs_SB.csv into R here:

```

**Your conclusion**:







## Question 3: NYC Flights

Do parts c), d), and e)

**Hint**: For part e), you may need to lean on Google. Google is a Data
Scientist's best friend; effective Googling is a skill that needs developping.
Example: Googling "DESIRED TASK dplyr" will return relevant results how to do a
desired task, but specific to the dplyr package.

```{r, echo=FALSE, message=FALSE}
library(nycflights13)
data(airports)
data(planes)
data(flights)
data(weather)
data(airlines)

# Do part c) here:


# Do part d) here:


# Do part e) here. Hint: 


```







## Question 4: Unisex Names... Revisited

```{r, echo=FALSE, message=FALSE}
library(babynames)

```



