
---
title: "Problem Set 05"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, October 21, 2016"
output:
  html_document
---




```{r, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)

# Packages containing data sets for today
library(okcupiddata)
library(gapminder)
```




## Question 1: Drug Use Amongst OkCupid Users

Let's revisit the 60,000 San Francisco OkCupid users in 2012 and consider the
variable `drug` reflecting users' answers to a question on their drug use.

```{r, echo=TRUE}
data(profiles)
# Convert as missing values in drugs to an actual category "N/A"
profiles <- profiles %>% 
  mutate(drugs=ifelse(is.na(drugs), "N/A", drugs))
```


#### a) 

Type in a series of commands that will output how many men and women there in this population.

**Solution:** This population skews male.

```{r, echo=TRUE}
# Write your code below:
male_vs_female <- profiles %>% 
  group_by(sex) %>% 
  summarise(count=n())
kable(male_vs_female)
```




#### b)

Create a visualization that shows the distribution of the different responses of
in variable `drugs`.

**Solution:**

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar()
```




#### c)

Create a visualization that shows the same information as in part a), but for
men and women separately. Who is more likely to never use drugs? Men or women?

**Solution:** We plot the raw counts of drug use split by sex. However, we can't 
compare these values directly because the population involves 35,829 males and
24,117 females as we saw above!

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar() +
  facet_wrap(~sex)
```

**Advanced**: So instead we must normalize to proportions. Note this was not 
expected of you for this problem set at the time, but we are now in a position 
to understand it. Here we `group_by` both `sex` and `drugs`, get counts, and 
compute proportions, we change the `group_by()` structure, and then instead of
summarizing, we mutate to compute a proportion. Ask yourself, what values sum to
1?

```{r, echo=TRUE}
proportions <- profiles %>% 
  group_by(sex, drugs) %>% 
  summarise(count = n()) %>% 
  group_by(sex) %>% 
  mutate(proportion = count/sum(count)) %>% 
  mutate(proportion = round(proportion, 2))
kable(proportions)
```

Now let's try to plot it. Recall from the [Lec12
Slides](https://rudeboybert.github.io/MATH116/assets/2-Data/barplots.html#8) 
there are two cases of barplots:

1. Where the counts are not pre-computed. what we've been using
1. Where the counts are pre-computed. This case.

We map the `proportion` to the `y` aesthetic. But for this to work, we need to
add a note to `geom_bar()` that we are specifying a `y` aesthetic by setting
`geom_bar(stat="identity")`

```{r, echo=TRUE}
ggplot(data=proportions, aes(x=drugs, y=proportion)) +
  geom_bar(stat="identity") +
  facet_wrap(~sex)
```

We now see that it is women who are more likely to self report that they never
use drugs.

#### d)

In lines 54-55 above we made sure to convert all missing values, encoded in R
as `NA`, to a specific response `N/A` i.e. not answered. Why do you think it was
important to do so?

**Solution:** Given the nature of the variable, some people might not be willing
to answer. It is important not to ignore these people or we might **bias** the
results.






## Question 2: Gapminder


#### a) 

Output a table that shows the mean and median GDP of all countries in the year 2007, but split by continent. Your table should be 5 rows and 3 columns

```{r}
data(gapminder)

mean_median_GDP_by_continent <- gapminder %>% 
  filter(year==2007) %>% 
  select(continent, gdpPercap) %>%
  group_by(continent) %>% 
  summarise(mean_GDP_per_capita = mean(gdpPercap), median_GDP_per_capita = median(gdpPercap))
kable(mean_median_GDP_by_continent)
```









## Question 3: Titanic

```{r, echo=TRUE}
data(Titanic)
# Convert the Titanic data to data frame format
Titanic <- Titanic %>% 
  as.data.frame()
```



#### a)

Output tables that compares survivor counts

1. between men and women
1. betweent the different classes


**Solution:** Let's first look at the `Titanic` data

```{r, echo=TRUE}
kable(Titanic)
```

We see that the variable `Freq` reports the **pre-computed** count of passengers
for each `Class`, `Sex`, `Age`, `Survived` category. The key difficulty of this
question was distiguising between using `n()` to count observations vs `sum()`.
Here, we need to `sum(Freq)`:

```{r, echo=TRUE}
survival_by_sex <- Titanic %>% 
  group_by(Sex, Survived) %>% 
  summarise(Freq=sum(Freq))
kable(survival_by_sex)

survival_by_class <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(Freq=sum(Freq))
kable(survival_by_class)
```



#### b)

For each comparison above, indicate who was most likely to survive.

* `Survival` by `Sex`: Female survival 344/(344+126) = 73.2%, Male survival 367/(367+1362) = 21.2%. Women were much more likely to survive.
* `Survival` by `Class`: 1st class was most likely to survive at 62.5%

**Advanced**: convert to proportions

```{r, echo=TRUE}
survival_by_class_prop <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(Freq=sum(Freq)) %>% 
  group_by(Class) %>% 
  mutate(proportion = Freq/sum(Freq)) %>% 
  mutate(proportion = round(proportion, 3))
kable(survival_by_class_prop)
```
