
---
title: "Problem Set 07"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 4, 2016"
output:
  html_document
---


## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 





## Instructions:

* Say you have your results in a data frame called `results`. Including
`kable(results)` in your code will print the table in your HTML page in a nice
and clean manner.


```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
```




## Question 1: "Women and Children First"

Do part c) as thoroughly as possible here:

```{r, echo=FALSE, message=FALSE}
data(Titanic)
# Convert Titanic data to tidy format
Titanic <- Titanic %>% 
  as.data.frame()

```

Did the "women-and-children" first policy hold? Argue using the results from the
above table.





## Question 2: "America Runs on Starbucks?"

Recreate the plot in part a) (with any improvements you think are warranted in
part c) in the code block. Don't worry about making the colors match.

```{r, echo=FALSE, message=FALSE}
# Copy and paste the line of code that loads DD_vs_SB.csv into R here:

```




## Question 3: "NYC Flights"

```{r, echo=FALSE, message=FALSE}
library(nycflights13)
data(airports)
data(planes)
data(flights)
data(weather)
data(airlines)

# Do part c) here:


# Do part d) here:


# Do part e) here:


```




## Question 4: "Unisex Names... Revisited"

```{r, echo=FALSE, message=FALSE}
library(babynames)

```



