
---
title: "Problem Set 08"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 11, 2016"
output:
  html_document
---



## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 



## Instructions/Hint:

* As always, please knit this first and read it over once.
* Interpreting probability:
    + **In general**: one interpretation of "the probability of x occuring" is 
    the proportion of the time `x` occurs across many, many, many
    attempts.
    + **Example**: "the probability of flipping a coin and getting heads is
    0.5 = 50%" can be interpreted as the proportion of heads occuring over many,
    many, many coin flips being one half.
    + For the purposes of this problem set, let "many, many, many times" mean 
    10,000 times.
* As always, I recommend you separate out the **what** vs the **how** by first 
sketching out a plan of what you are going to do for each question on paper.

```{r, message=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)

# For printing clean tables using the kable() function
library(knitr)

# Set random number generator seed value:
set.seed(76)
```



## Question 1:

We will be conducting the study from Lec21 on Friday November 4th on whether or
not wearing shoes while sleeping causes people to wake up with headaches.

#### a)

Dr. Quack goes through the medical files for 1000 of their patients and obtains 
variables:

* `shoes_on`: whether or not the patient slept with their shoes on
* `headache`: whether or not the patient woke up with a headache

where in both cases `0` indicates false and `1` indicates true. Dr. Quack 
presents their findings in a data frame `study_data_1` which is loaded from the
web below. Questions:

1. Who had a higher headache rate: students who slept with their shoes on or those
who didn't? Is this difference significant?
1. Can you prove or disprove the claim that sleeping with shoes on causes people
to wake up with a headache with the information below? Why or why not?

```{r, echo=TRUE, message=FALSE}
study_data_1 <- readr::read_csv("https://rudeboybert.github.io/MATH116/assets/PS/study_data_1.csv")

```



#### b)

Take your code for part a), but change it so that retroactively we had data from
a **randomized experiment**. Questions:

* Who now had a higher headache rate: students who slept with their shoes on or those
who didn't? Is this difference significant?
* Can you prove or disprove the claim that sleeping with shoes on causes people
to wake up with a headache with the information below? Why or why not?


```{r, echo=TRUE, message=FALSE}

```


#### c)

Now say you also have access to information as to whether or not the patient got
wasted the night before. Generate a table that compares headache rates for those
patients who slept with shoes on versus those who didn't **controlling** for 
whether or not they got wasted, in other words, taking into account whether or
not they got wasted.

```{r, echo=TRUE, message=FALSE}
study_data_2 <- readr::read_csv("https://rudeboybert.github.io/MATH116/assets/PS/study_data_2.csv")

```




## Question 2:

You are going to compute via simulation and not via mathematical formulas, the 
**probability distribution** of all possible sums of two die rolls: for all 
possible sums of rolling two dice (x=2, x=3, ..., x=12), we compute
Probability(sum of rolling two dice = x).

1. Write code that will simulate die rolls and manipulate the output so that
you end up with a data frame `probs` with 11 rows and 2 columns:
    + `sum`: each value of the possible sum of rolling two dice
    + `probability`: the observed probability of each value of `sum`
1. Once you've fully created it, change the `eval=FALSE` to `eval=TRUE` in the 
starting line of the code block below to have the plot get generated.
1. Note even though `sum` is a numerical variable, we'll treat it as a 
categorical variable by using the `factor()` command and using a bar chart. It 
looks cleaner this way.

```{r, eval=FALSE}
# Write your code to create the data frame probs here:


# Suggested ggplot call. Feel free to change it:
ggplot(probs, aes(x=factor(sum), y=probability)) +
  geom_bar(stat="identity") +
  labs(x="Sum", y="Probability of Sum", title="Outcomes of Rolling Two Dice")
```




## Question 3:

Problem 3.20 on page 162 of the OpenIntro Textbook on "With and without 
replacement". For this question, find all probabilities using R's sampling 
capabilities.

Part d) is the only required part for this problem set. Parts a), b), and c) are 
bonus parts.

#### a) and b) from text

**Starter Code**: We define a `room` and `stadium`, both half female, but with 
populations of size `n=10` and `n=10000`. Let `1` represent a female and `0` 
represent a male, so that to count the number of females we need to only `sum()`
a series of `0`'s and `1`'s.

```{r}
room <- rep(c(1, 0), each=5)
stadium <- rep(c(1, 0), each=5000)

```

Fill in the table with the results from your simulations in the code block below

Sampling Type        | Room   | Stadium
-------------------- | -------| -------
With replacement     | 0.000  | 0.000
Without replacement  | 0.000  | 0.000


#### c) from text


#### d)

For the stadium case, answer the following questions about the components of the
"powerball" analogy:

**Atrributes of the Lottery Machine**:

* How many balls do you have?
* What are written on the balls?
* Do the balls have equal probability of being picked?


**Attributes of the Drawing**:

* How are you drawing the balls?
* How many balls do you draw? 
* What are you recording about each drawn ball?




