
---
title: "Problem Set 06"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, October 28, 2016"
output:
  html_document
---


## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 



## Instructions:

* As I said in lab on Thursday, unless you are **extremely** comfortable with
the tools of `ggplot2` and `dplyr`, do not attempt to answer the questions below
immediately in this R Markdown file. Rather, focus on separating the
    + **"What"** you are going to do: the blueprint/plan
    + **"How"** you are going to do it:  `ggplot2` and `dplyr` coding
* Confusing the "what" and the "how" will only make this task more difficult.
* For example, the "what" step could be tackled as follows:
    + Draw on paper, a blackboard, or a whiteboard what your starting data frames
    look like and
        1. what your resulting data frames should look like
        1. what any visualizations should look like
    + Write out in 
    [pseudocode](http://image.slidesharecdn.com/pseudocode-111128151539-phpapp02/95/introduction-to-pseudocode-38-638.jpg?cb=1393513763)
    what your R code will look like. Pseudocode is informal and rough code that 
    doesn't necessarily need to work, but still illustrates what you want to do.
    More than anything else, writing pseudocode out forces you to think your
    plan out loud.
* Only after you a plan for the "what" should you start the "how" i.e. coding.
* I **HIGHLY** encourage you to work in groups on this problem set, on both the
"what" and "how" stages.




## Question: `nycflights13` data

```{r, echo=FALSE, message=FALSE}
# Load all pacakges necessary
library(ggplot2)
library(dplyr)
library(knitr)

# Load all data sets in nycflights
library(nycflights13)
data(flights)
data(planes)
data(airlines)
data(weather)
data(airports)
```

You are a junior analyst in the airline industry, specifically working for 
Virgin America. You are tasked with calculating Virgin America's [available seat
miles](https://en.wikipedia.org/wiki/Available_seat_miles) for all flights 
leaving each of the three New York Metropolitan area airports separately in 
2013, and comparing these figures with those of ExpressJet Airlines Inc, a close
rival to your company. Available seat miles are the fundamental units of
production for a passenger-carrying airline and it is a measure of capacity. You
are asked for two deliverables:

1. Present these results in
    + A well-formatted and easy to read table
    + A cleanly formatted visualization
1. Give a two sentence executive summary on the comparison in available seat
miles between Virgin America and ExpressJet Airlines Inc for the three NYC
airports.


Management offers the following helpful tips:

* Keep `View()`ing your data
* Figure 5.8 in [Chapter
5.3](https://rudeboybert.github.io/IntroStatDataSciences/5-manip.html#joiningmerging-data-frames)
gives a sense of what information you have to work with.
* The help files contain the documentation. For example `?flights`
* Say you have your results in a data frame called `results`. Including `kable(results)` in your code
will print the table in your HTML page in a nice and clean manner.
* You can `group_by` more than one variable. Example say you had another data
set that you want to group by `state` and `city`, you would use `group_by(state,
city)`.


```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}

```
