
---
title: "Problem Set 07"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 4, 2016"
output:
  html_document:
    df_print: kable
---



```{r, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
```



## Comment

This was a difficult problem set.




## Question 1: Women and Children First

**Research question**: Did the "women-and-children" first
policy hold? Hint: the answer is yes.

Argue your conclusion using the results from an outputed table. Do as much as
you can computationally to minimize the number of hand calculations you may need
to make (if any).

```{r, echo=FALSE, message=FALSE, echo=TRUE}
data(Titanic)
# Convert Titanic data to tidy format
Titanic <- Titanic %>% 
  as.data.frame()

Titanic %>% 
  group_by(Survived, Sex, Age) %>% 
  summarise(Freq=sum(Freq)) 
```


**Conclusion**: 

* `338/(338+1329)` = 20.3% of male adults survived
* `(29+28+316)/(29+28+316+35+17+109)` = 69.8% of everyone else survived

Don't need no PhD in statistics to know this difference is **statistically significant** i.e.

1. Survival wasn't random
1. There was a mechanism dictating who had greater chances of survival







## Question 2: America Runs on Starbucks?

**Research question**: Do Starbucks tend to locate in higher income neighbors,
while Dunkin Donuts does not i.e. they tend to locate in lower income
neighborhoods." 

**Your conclusion**: Here is the plot from the midterm (without colors).

```{r, echo=FALSE, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
DD_vs_SB <- read.csv("DD_vs_SB.csv")

ggplot(DD_vs_SB, aes(x=median_income, y=shops_per_1000)) +
  geom_point() + 
  facet_wrap(~Type) +
  geom_smooth(method="lm", se=FALSE) + 
  labs(x="Median Household Income", y="# per 1000 people", title="Coffee/Cafe Comparison in Eastern MA")
```

I'm going to do three things to improve it:

* Use a smoother that isn't a regression line
* Change the y-axis scale to where most of the points are: between 0 and 0.75
* Use `alpha=0.1` in the `geom_point()` to try to see better what's going on at `y=0`

```{r, echo=FALSE, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
ggplot(DD_vs_SB, aes(x=median_income, y=shops_per_1000)) +
  geom_point(alpha=0.1) + 
  facet_wrap(~Type) +
  geom_smooth(se=FALSE) + 
  labs(x="Median Household Income", y="# per 1000 people", title="Coffee/Cafe Comparison in Eastern MA") +
  coord_cartesian(ylim=c(0, 0.75))
```

**Advanced**: Why did I use `coord_cartesian(ylim=c(0, 0.75))` and not 
`ylim(c(0, 0.75))`? Because of **point clipping**. See the [ggplot2 
cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf),
back page, bottom right corner, Zooming section to see the difference in
behavior. Some times you want to "clip" points and others you don't. 

In our case, if you clip points, it will alter the behavior of the smoother
since it will be based on fewer points. This also relevant to Question 3.d)
below.





## Question 3: NYC Flights

```{r, echo=FALSE, message=FALSE, echo=TRUE}
library(nycflights13)
data(airports)
data(planes)
data(flights)
data(weather)
data(airlines)
```



#### c)

Write the rough pseudocode that will output a table presenting the median
departure delay of all flights for each airline leaving Newark (airport code
`EWR`) in reverse alphabetical order.

**Solution**: This gets the job done:

```{r, echo=FALSE, message=FALSE, echo=TRUE}
median_delays_by_carrier_EWR <- flights %>% 
  filter(origin == "EWR") %>% 
  group_by(carrier) %>% 
  summarise(med_delay = median(dep_delay, na.rm=TRUE)) %>% 
  arrange(desc(carrier))

median_delays_by_carrier_EWR
```

**Advanced**: But this is a little more elegant: using the `name` variable in the `airlines`
data set instead of `carrier` codes.

```{r, echo=FALSE, message=FALSE, echo=TRUE}
flights %>% 
  # join the flights to airlines data sets here!
  left_join(airlines, by="carrier") %>% 
  filter(origin == "EWR") %>% 
  group_by(name) %>% 
  summarise(med_delay = median(dep_delay, na.rm=TRUE)) %>% 
  arrange(desc(name))
```



#### d)

Name a graphic that would best show all the information contained in the
tables in part b)/c).

**Solution**: This gets the job done:

```{r, echo=FALSE, message=FALSE, echo=TRUE, fig.width=16/2, fig.height=9/2}
EWR_flights <- flights %>% 
  filter(origin == "EWR")

ggplot(data=EWR_flights, aes(x=carrier, y=dep_delay)) +
  geom_boxplot()
```

**Advanced**: Let's zoom 



#### e)

Write the rough pseudocode that will tabulate the mean humidity level 
recorded for all flights leaving New York City in July 2013.

**Solution**:

The tough part here was recognizing that 






## Question 4: Unisex Names... Revisited

```{r, echo=FALSE, message=FALSE}
library(babynames)

```



