
---
title: "Problem Set 09"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 18, 2016"
output:
  html_document
---

Investigate whether economics majors did better than all other majors on the
final exam for my intro stats class back in the day. Tweak the code from
`Lec25.R`.

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)

# Set see value
set.seed(76)

# Copy the code that loads grades.csv into RStudio here.
# Note you will need to change this line to whatever loads the data for you:
grades <- read_csv("../PS/grades.csv")
```



## Exploratory Data Analysis

It is not immediately apparent if one did better:

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
ggplot(data=grades, aes(x=major, y=final)) +
  geom_boxplot()
ggplot(data=grades, aes(x=final)) +
  geom_histogram(binwidth = 0.25) + 
  facet_wrap(~major)
```

**Important**: We have to be very careful to note if we are subtracting:

* `Other-Economics`
* `Economics-Other`

and be consistent. It seems we are doing `Other-Economics`: 

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
mean(final ~ major, data=grades)
mean(final ~ major, data=grades) %>% diff()
```

Econ majors did about 7.8% better on average.

## Hypothesis Test

We proceed with the hypothesis test

* $H_0$: econ and other majors did equally well i.e. no difference
* $H_A$: i.e. there is a difference

Recall that under $H_0$, since there is no difference between the two groups,
the variable `major` is meaningless, so we can permute/shuffle it around and it
doesn't matter:

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2, cache=TRUE}
# Compute observed test statistic
observed_diff <- mean(final ~ major, data=grades) %>% diff()

# Simulate the null distribution of the test statistic:
simulations <- do(10000) * mean(final ~ shuffle(major), data=grades)
simulations <- simulations %>%
  as_data_frame() %>% 
  mutate(difference=Other-Economics)

# Compare what we observed (red) to what happens if we assume no difference:
ggplot(data=simulations , aes(x=difference)) +
  geom_histogram() +
  geom_vline(xintercept = observed_diff, col="red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  labs(x="Other Avg - Econ Avg")
```


## Conclusion

IMO I wouldn't say Econ did significantly better. We fail to reject $H_0$ that
they did equally well.  i.e. Maybe one group is in fact better than the other, but
we don't have evidence to suggest that here.

**Note**: this was an example of a *two-sided two-sample permutation test for differences in means*

1. two-sided: type of alternative hypothesis
1. two-sample: i.e there are two groups
1. permutation test: we constructed the null distribution via permutations/simulations
1. difference in means: comparing means

More on these options on Monday.



## p-Value

The p-value is the probability of observing a test statistic just as or more 
extreme than the one observed.  In our case since $H_A: \mu_{other} - \mu_{econ} \neq 0$: i.e. there is a difference, we can have either

* A more extreme negative difference 
* A more extreme positive difference

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2, cache=TRUE}
simulations %>%
  mutate(more_extreme_left = difference <= observed_diff) %>% 
  summarise(more_extreme_left = sum(more_extreme_left))
```

There are 1181 values less than the observed difference of -7.8%. So the p-value is 

$$
\frac{2 \times 1181 + 1}{10000 + 1} = 0.236
$$
**Notes**:

1. Note we doubled the 1181 because we have a two-sided alternative $H_A: \mu_{other} - \mu_{econ} \neq 0$
1. We add 1 to both the numerator and denominator since we also need to account for the observed test statistic itself



## Other Alternatives

* Say we had $H_A: \mu_{other} - \mu_{econ} < 0$, more extreme would mean more
negative, i.e. more to the left of the observed test statistic, i.e. the p-value
would be
$$
\frac{1181 + 1}{10000 + 1} = 0.118
$$
* Say we had $H_A: \mu_{other} - \mu_{econ} > 0$, more extreme would mean more
positive, i.e. more to the right of the observed test statistic, i.e. the p-value
would be
$$
\frac{(10000-1181) + 1}{10000 + 1} = 0.882
$$
