
---
title: "Problem Set 09"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, November 18, 2016"
output:
  html_document
---



## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 



```{r, message=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)
```



## Instructions/Hints:

* Only include that is absolutely necessary to conducting the test. For example:
    + Remove all `View()`'s since R Markdown will generate an error
    + Do not include the sidenotes from Lec25.R. They were included only to assist your learning
* If you feel up to it, compute the p-value!


## Question 1:

Investigate whether economics majors did better than all other majors on the
final exam for my intro stats class back in the day. Tweak the code from
`Lec25.R`.


```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
# Copy the code that loads grades.csv into RStudio here:


# Do your exploratory data analysis here:


# Perform your hypothesis test here:


```


What is your conclusion?

