
---
title: "Problem Set 11"
author: "WRITE YOUR NAME HERE"
date: "Due Monday, December 9, 2016 at 11am"
output:
  html_document
---



## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 



```{r, message=FALSE, echo=FALSE}
# Load packages:
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)
library(knitr)
```



## Instructions/Hints:

* Knit this once and read over all questions
* Set the seed value on line 69 to be something other than 76
* Fill in this Google Sheet with your answers



## Scenario:

You are charged to study the average age of members of congress (House of
Representatives and Senate) at the start of their term for all sessions of
Congress between:

* The [80th United States Congress](https://en.wikipedia.org/wiki/80th_United_States_Congress) 
that started on January 3, 1947
* The [113th United States Congress](https://en.wikipedia.org/wiki/113th_United_States_Congress) 
that started on January 3, 2013

In particular, you are interested in answering the following three questions:

1. What is the average age for **all** members of Congress between the 80th and
113th sessions of Congress? Is it 30?
1. Comparing just the 113th and the 80th sessions of Congress, are members of
congress on average younger now (in 2013) or then (in 1947)?
1. Considering just the 113th session on Congress, who on average are younger?
Democrats or Republicans?

However, researching birthdays and age at the start of the terms for all 18635
members of congress between the 80th and 113th sessions of Congress is painstakingly
boring work. So you decide to use sampling. Run the following two lines: 

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
source("https://rudeboybert.github.io/MATH116/assets/PS/raw_data/sampling.R")
# Change this seed value to be your favorite number:
seed_value <- 76
```



## Question 1: CI for Population Mean

Run these two lines to get your sample over all members of congress between the
80th and 113th sessions of Congress.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
set.seed(seed_value)
congress <- get_sample_of_congress(n = 100)
```


#### a) EDA

Create one plot of exploratory data analysis.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```


#### b) 95% CI with `n=100`

Just as in the LC for Lec31, create a 95% confidence interval for $\mu$, the
true population mean age of all members of congress between the 80th and 113th
sessions of Congress.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```

**Your Answers**: 

* My confidence interval is: ( , )
* My answer to the first question above is:


#### c) 95% CI with `n=1000`

Repeat part b), but now taking a random sample of 1000 members of congress.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
set.seed(seed_value)
congress <- get_sample_of_congress(n = 1000)
```

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```

**Your Answers**: 

* My confidence interval is: ( , )
* My answer to the first question above is:







## Question 2: CI for Difference of Population Means

Run these two lines to get your random sample of

* `n_80` 40 members of congress from the 80th session
* `n_113` 35 members of congress from the 113th session

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
set.seed(seed_value)
congress_80_and_113 <- get_sample_of_congress_80_and_113(n_80 = 40, n_113 = 35)
```


#### a) EDA

Create one plot of exploratory data analysis.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```


#### b) 95% CI

Create a 95% confidence interval for $\mu_{80}-\mu_{113}$, the true **population
difference in means** of the average age in the 80th session minus the average 
age in the 113th session. 

**Hint**: The formula to approximate the standard error for a difference of means 
$\overline{x}_1 - \overline{x}_2$ is

$$
SE_{\overline{x}_1 - \overline{x}_2} = \sqrt{ 
\left( \frac{(n_1-1)s_1^2 + (n_2-1)s_2^2}{n_1+n_2-2} \right) 
\left( \frac{1}{n_1} + \frac{1}{n_2} \right) 
}
$$

Nasty! Why don't I just give it to you:

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
SE <- SE_diff_means(congress_80_and_113, variable="age", group_by_variable = "congress")
```

**Your Answers**: 

* My confidence interval is: ( , )
* My answer to the second question above is:







## (BONUS) Question 3: Difference of Population Means

Answer the third question above based on a sample of size 544. Don't forget to
do an EDA!

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
set.seed(seed_value)
congress_113 <- get_sample_of_congress_113(n_113 = 544)
```

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```

**Your Answers**: 

* My answer to the third question above is:

