
---
title: "Problem Set 10"
author: "WRITE YOUR NAME HERE"
date: "Due Monday, November 28, 2016 at 5pm"
output:
  html_document
---



## Please Indicate

* Who you collaborated with. If no one, please indicate so: 
* Approximately how much time did you spend on this problem set: 
* What, if anything, gave you the most trouble: 



```{r, message=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)
```



## Instructions/Hints:

* 





## Question 1:

Answer the question from Question 3 from the midterm "The Dude Rolling Die": is he 
full of BS or can he actually correctly predict die rolls?

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}

```





## Question 2:

#### a) 

Let's consider the hiring data in `hiring.csv`

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
# Copy the code that loads hiring.csv into RStudio here:


# Do your exploratory data analysis here:


# Perform your hypothesis test, including p-value


```


#### b) 

Who penalized women applicants more on average? The male evaluators or the female evaluators?

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}



```





## Question 3:

Proporportions fail to reject.



