---
title: "Interesting Study"
author: "Albert Y. Kim"
date: "Fri Nov 18, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
slides > slide:not(.nobackground):after {
  content: ""
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(mosaic)
library(knitr)
```



## Effect of Gender

What is the effect of gender on being hired for a job?

> * One way to identify causal effect is to randomly assign individuals to treatment and control groups
> * Our case: to identify effect of gender on hireability score, you need to randomly assign gender
> * Randomly assign gender? Say what?





## Study at Yale

> * Scientists were presented with application materials from a "student" applying for a lab manager position
> * Name at the top of resume was randomly chosen between
>     + Very male sounding name
>     + Very female sounding name






## Data

```{r, echo=FALSE, message=FALSE}
hiring <- read_csv("~/Documents/Teaching/MATH116/116-Webpage/assets/PS/hiring.csv")
set.seed(79)
hiring %>% 
  select(applicant_gender, hireable) %>%
  mutate(hireable=round(hireable, 3)) %>% 
  sample_n(6) %>% 
  kable()
```



## Question

Despite only gender of name at the top of application being different, who do you think
got higher hireability scores?



## Twist

> * Researchers didn't only consider the gender of applicant, but also gender of **evaluator**!
> * Now we can compare male vs female applicant scores **controlling for** gender of evaluator.


## Data

```{r, echo=FALSE, message=FALSE}
hiring <- read_csv("~/Documents/Teaching/MATH116/116-Webpage/assets/PS/hiring.csv")
set.seed(79)
hiring %>% 
  select(applicant_gender, evaluator_gender, hireable) %>%
  mutate(hireable=round(hireable, 3)) %>% 
  sample_n(6) %>% 
  kable()
```


## Question

Who do you think was harsher on the female applicants?

* Male evaluators
* Female evaluators









