
---
title: "Problem Set 09"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, April 21, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
---


Load necessary packages:

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(nycflights13)
set.seed(76)
```



## Question 1:

### Code

Create the visualization from Question 3.b) of Midterm II:

**Answer**: 

* We only care about the 4 aces out of 52. Let `1` denote the 4 aces, and `0` denote the remaining
48 cards. Note `rep()` is a command to repeat a certain value the specified number of times.
* We sample 5 cards *without* replacement

We plot the results:

```{r}
deck <- c(rep(1, times=4), rep(0, times=48))
simulation <- do(10000) * resample(deck, size=5, replace=FALSE)
simulation <- simulation %>% 
  mutate(num_aces = V1 + V2 + V3 + V4 + V5)
ggplot(simulation, aes(x=num_aces)) +
  geom_bar()
```


```{r}
simulation %>% 
  group_by(num_aces) %>% 
  summarise(count=n())
```







### Analysis

1. What is the probability of getting a pair of aces? 
1. What is the probability of getting one or more aces?

**Answer**:

1. This occured 368 out of 10000 times, so roughly 0.0368 = 3.68% of the time.
1. No aces occured 6601 out of 10000 times, so roughly 0.6601 = 66.01%. This getting 1 or more is
1-0.6601=0.3399 = 33.99%.


**Note**:

* How rare getting 4 aces is! If you're playing poker and get 4 aces, bet your house!
* This question can also be solved mathematically using the [hypergeometric
distribution](https://en.wikipedia.org/wiki/Hypergeometric_distribution), which
gives the *exact* quantity we could repeat the above experiment infinitely. So
the probability of picking `x=2` aces out of `k=5` cards sampled without
replacement, from a deck with `m=4` total aces and `n=48` other cards is
*exactly*:

```{r}
dhyper(x=2, m=4, n=48, k=5)
```

Our simulated answer is pretty close from part 1!



## Question 2:

### Code

Do Question 2.d) from Midterm II 

1. by joining `flights` and `weather` by year, month, day, hour
1. by joining `flights` and `weather` by year, month, day, hour, and origin as
per the diagram at the start of Question 2.

**Hint**: Google is a programmer's best friend.

**Answer**: In order to join by multiple variables, we use the `c()` for
"concatenate" function in the `by` argument:

```{r}
flights %>% 
  left_join(weather, by=c("year", "month", "day", "hour")) %>% 
  filter(month==7) %>% 
  summarise(mean_humid = mean(humid, na.rm=TRUE))

flights %>% 
  left_join(weather, by=c("year", "month", "day", "hour", "origin")) %>% 
  filter(month==7) %>% 
  summarise(mean_humid = mean(humid, na.rm=TRUE))
```


### Analysis

Why are the answers different?

**Answer**: Because the data are linked for a specific time *for each airport*. 
For example the temperature at Newark may be different than at JFK. If we don't 
make it absolutely clear that the join needs to happen along `origin` as well,
we may end up joining weather at a time at one airport with flight information
for the same time *but at a different airport*.
 
