
---
title: "Problem Set 07"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, April 7, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
    df_print: kable
---


## Question: Available Seat Miles

```{r, echo=TRUE, message=FALSE}
# Load all pacakges necessary
library(ggplot2)
library(dplyr)
library(knitr)

# Load all data sets in nycflights
library(nycflights13)
data(flights)
data(planes)
data(airlines)
data(weather)
data(airports)
```

You are a junior analyst in the airline industry, specifically working for 
Virgin America. You are tasked with calculating Virgin America's total
[available seat miles](https://en.wikipedia.org/wiki/Available_seat_miles) for
all flights leaving each of the three New York Metropolitan area airports
separately, and comparing these figures with those of ExpressJet Airlines Inc, a
close rival to your company. Available seat miles are the fundamental units of 
production for a passenger-carrying airline; it is a measure of **capacity**. 

1. You are asked for two deliverables: Present these results in
    1. A well-formatted and easy to read table
    1. A cleanly formatted visualization
1. Give a two sentence "executive summary" on the comparison in available seat
miles between Virgin America and ExpressJet Airlines Inc for the three NYC
airports.

Please heed the "tao of data analysis" [tips](https://rudeboybert.github.io/MATH116/notes.html#lec18_-_fri_324:_the_tao_of_data_analysis)
before starting.


## Deliverable 1: Table

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Write your R code below:

```




## Deliverable 2: Visualization

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Write your R code below:

```



## Executive Summary

Write your two sentence executive summary here.

