
---
title: "Problem Set 08"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, April 14, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
    df_print: kable
---


Load necessary packages:

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(knitr)
set.seed(76)
```



## Question 1:

You are going to compute via simulation and not via mathematical formulas, the 
**probability distribution** of all possible sums of two die rolls: for all 
possible sums of rolling two dice (x=2, x=3, ..., x=12), we compute
Probability(sum of rolling two dice = x).

1. Write code that will simulate die rolls and manipulate the output so that
you end up with a data frame `probs` with 11 rows and 2 columns:
    + `sum`: each value of the possible sum of rolling two dice
    + `probability`: the observed probability of each value of `sum`
1. Once you've fully created it, change the `eval=FALSE` to `eval=TRUE` in the 
starting line of the code block below to have the plot get generated.
1. Note even though `sum` is a numerical variable, we'll treat it as a 
categorical variable by using the `as.factor()` command and using a bar chart. It 
looks cleaner this way.


**Solution**:

Key points to understand:

* We first simulate sums of two die rolls. These are the possible outcomes: 2, 3, 4, ..., 11 or 12.
* We then count the number of times each outcome occurred using `n()` and not `sum()`
* We convert the counts into to proportion by dividing by the number of simulations we did: 10000


```{r, eval=TRUE}
# Write your code to create the data frame probs here:
die <- c(1:6)
two_die_rolls <- do(10000) * resample(die, size=2, replace = TRUE)

probs <- two_die_rolls %>% 
  mutate(sum=V1+V2) %>% 
  group_by(sum) %>% 
  summarise(count=n()) %>% 
  mutate(probability=count/10000)

# Suggested ggplot call. Feel free to change it:
ggplot(probs, aes(x=as.factor(sum), y=probability)) +
  geom_bar(stat="identity") +
  labs(x="Sum", y="Probability of Sum", title="Outcomes of Rolling Two Dice")
```




## Question 2:

Problem 3.20 on page 162 of the OpenIntro Textbook on "With and without 
replacement". For this question, find all probabilities using R's sampling 
capabilities.


**Solution**:

#### Advanced: a) and b) from text

**Starter Code**: We define a `room` and `stadium`, both half female, but with 
populations of size `n=10` and `n=10000`. Let `1` represent a female and `0` 
represent a male, so that to count the number of females we need to only `sum()`
a series of `0`'s and `1`'s.

```{r}
room <- rep(c(1, 0), each=5)
stadium <- rep(c(1, 0), each=5000)

# Sampling with and without replacement from the room
room_sample_with_replace <- do(10000)*resample(room, size=2, replace=TRUE)
room_sample_with_replace %>% 
  mutate(num_female = V1+V2) %>% 
  group_by(num_female) %>% 
  summarise(count=n())

room_sample_without_replace <- do(10000)*resample(room, size=2, replace=FALSE)
room_sample_without_replace %>% 
  mutate(num_female = V1+V2) %>% 
  group_by(num_female) %>% 
  summarise(count=n())

# Sampling with and without replacement from the stadium
stadium_sample_with_replace <- do(10000)*resample(stadium, size=2, replace=TRUE)
stadium_sample_with_replace %>% 
  mutate(num_female = V1+V2) %>% 
  group_by(num_female) %>% 
  summarise(count=n())

stadium_sample_with_replace <- do(10000)*resample(stadium, size=2, replace=FALSE)
stadium_sample_with_replace %>% 
  mutate(num_female = V1+V2) %>% 
  group_by(num_female) %>% 
  summarise(count=n())
```

Fill in the table with the results from your simulations in the code block below

Sampling Type        | Room   | Stadium
-------------------- | -------| -------
With replacement     | 0.250  | 0.2485
Without replacement  | 0.223  | 0.2501

**Conclusion**:

* The effect of sampling with vs without replacement in the room leads to a difference of 2.7%
* The effect of sampling with vs without replacement in the stadium leads to a difference of 0.0016%


#### Advanced: c) from text

**Moral of the Story**: The larger the population, the less the difference in results between sampling
with vs without replacement matter.






