
---
title: "Problem Set 06"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, March 24, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
    df_print: kable
---

Load necessary packages:

```{r, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
```




## Question 1: Drug Use Amongst OkCupid Users

Let's revisit the 60,000 San Francisco OkCupid users in 2012 and consider the 
variable `drug` reflecting users' answers to a question on their drug use. Run
this code first:

```{r, echo=TRUE}
library(okcupiddata)
data(profiles)
# Convert as missing values in drugs to an actual category "N/A"
profiles <- profiles %>% 
  mutate(drugs = ifelse(is.na(drugs), "N/A", drugs))
```


**a)** Type in a series of commands that will output a table of how many men and
women there are in this population.

**Solution:** This population skews male.

```{r, echo=TRUE}
# Write your code below:
male_vs_female <- profiles %>% 
  group_by(sex) %>% 
  summarise(count=n())
male_vs_female
```


**b)** Create a visualization that shows the distribution of the different
responses of in variable `drugs`.

**Solution:**

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar()
```

**c)** Create a visualization that shows the same information as in part a), but
for men and women separately. Who is more likely to say they never use drugs?
Men or women?

**Solution:** We plot the raw counts of drug use split by sex. However, we can't 
compare these values directly because the population involves 35,829 males and
24,117 females as we saw above!

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar() +
  facet_wrap(~sex)
```

**Advanced**: So instead we must normalize to proportions. Note this was not 
expected of you for this problem set at the time, but we are now in a position 
to understand it. Here we `group_by` both `sex` and `drugs`, get counts, and 
compute proportions, we change the `group_by()` structure, and then instead of
summarizing, we mutate to compute a proportion. Ask yourself, what values sum to
1?

```{r, echo=TRUE}
proportions <- profiles %>% 
  group_by(sex, drugs) %>% 
  summarise(count = n()) %>% 
  group_by(sex) %>% 
  mutate(proportion = count/sum(count)) %>% 
  mutate(proportion = round(proportion, 2))
proportions
```

Now let's try to plot it. Recall when we covered barplots, we saw there are two cases of barplots:

1. Where the counts are not pre-computed. what we've been using
1. Where the counts are pre-computed. This case.

We map the `proportion` to the `y` aesthetic. But for this to work, we need to
add a note to `geom_bar()` that we are specifying a `y` aesthetic by setting
`geom_bar(stat="identity")`

```{r, echo=TRUE}
ggplot(data=proportions, aes(x=drugs, y=proportion)) +
  geom_bar(stat="identity") +
  facet_wrap(~sex)
```

We now see that it is women who are more likely to self report that they never
use drugs.

**d)** After we loaded the data above via `data(profiles)`, we made sure to
convert all missing values, encoded in R as `NA`, to a specific response `N/A`
i.e. not answered. Why do you think it was important to do so?

**Solution:** Given the nature of the variable, some people might not be willing
to answer. It is important not to ignore these people or we might **bias** the
results.






## Question 2: Gapminder

We're going revisit the Gapminder data from PS-05. Run this code first:

```{r}
library(gapminder)
data(gapminder)
```


**a)** Output a table that shows the mean and median GDP of all countries in the
year 2007, but split by continent. Your table should be 5 rows and 3 columns.


```{r}
mean_median_GDP_by_continent <- gapminder %>% 
  filter(year==2007) %>% 
  select(continent, gdpPercap) %>%
  group_by(continent) %>% 
  summarise(
    mean_GDP_per_capita = mean(gdpPercap),
    median_GDP_per_capita = median(gdpPercap)
  )
mean_median_GDP_by_continent
```









## Question 3: Titanic

Let's study data relating to who survived the Titanic disaster. Run this code
first:

```{r, echo=TRUE}
data(Titanic)
# Convert the Titanic data to data frame format
Titanic <- Titanic %>% 
  tbl_df()
```


**a)** Output tables that compares survivor counts

1. between men and women
1. betweent the different classes


**Solution:** Let's first look at the `Titanic` data

```{r, echo=TRUE}
Titanic
```

We see that the variable `n` reports the **pre-computed** count of passengers
for each `Class`, `Sex`, `Age`, `Survived` category. The key difficulty of this
question was distiguising between using `n()` to count observations vs `sum()`.
Here, we need to `sum(n)`:

```{r, echo=TRUE}
survival_by_sex <- Titanic %>% 
  group_by(Sex, Survived) %>% 
  summarise(n=sum(n))
survival_by_sex

survival_by_class <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(n=sum(n))
survival_by_class
```



**b)** For each comparison above, indicate who was most likely to survive.

* `Survival` by `Sex`: Female survival 344/(344+126) = 73.2%, Male survival 367/(367+1362) = 21.2%. Women were much more likely to survive.
* `Survival` by `Class`: 1st class was most likely to survive at 62.5%

**Advanced**: convert to proportions

```{r, echo=TRUE}
survival_by_class_prop <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(n=sum(n)) %>% 
  group_by(Class) %>% 
  mutate(proportion = n/sum(n)) %>% 
  mutate(proportion = round(proportion, 3))
survival_by_class_prop
```
