
---
title: "Problem Set 10"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, April 28, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
---

Load necessary packages:

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)
set.seed(76)
```



## Question 1:

Answer the question "Did Econ majors do better on the intro stats final than
non-Econ majors?" using the `grades.csv` data set from Lec28. Note: save the
file `grades.csv` in the same folder/directory as `PS-10.Rmd`.


#### Hypothesis Test Setup

We have

* $H_0: \mu_E - \mu_O = 0$ vs
* $H_A: \mu_E - \mu_O > 0$

or equivalently

* $H_0: \mu_O - \mu_E = 0$ vs
* $H_A: \mu_O - \mu_E < 0$

Note we are dealing with a  *one-sided* alternative!


#### Observed Test Statistic

First, we compute the observed test statistic, noting carefully that it's the
mean of the other students minus the mean of the econ students: 
$\overline{x}_O - \overline{x}_E$:

```{r, message=FALSE}
grades <- read_csv("grades.csv")
mean(final ~ major, data=grades)
mean(final ~ major, data=grades) %>% diff()
observed_diff <- mean(final ~ major, data=grades) %>% diff()
```

It seems that Econ majors did 7.76% better, but is this difference **statistically** significant?

#### Null Distribution

Next, we construct the null distribution of the difference in means *assuming* there
is no true difference. Note when computing `difference`, we make sure that we
subtract in the correct order:

```{r, message=FALSE}
simulations <- do(10000) * mean(final ~ shuffle(major), data=grades)
simulations <- simulations %>%
  mutate(difference=Other-Economics)
```

#### Comparison

Finally, we compare the

* observed test statistic
* null distribution of the test statistic

```{r, message=FALSE}
ggplot(data=simulations , aes(x=difference)) +
  geom_histogram() +
  labs(x="Avg of Others - Avg of Economics") +
  geom_vline(xintercept = observed_diff, col="red")
```

The observed difference of -0.0776 = -7.76% doesn't seem all that extreme. i.e.
it seems plausible 

#### p-Value

In this case, since $H_A: \mu_O - \mu_E < 0$, more extreme means *more negative*!

So the p-value is the proportion of time

```{r, message=FALSE}
simulations %>% 
  mutate(more_extreme = difference < -0.0776) %>% 
  summarise(n_more_extreme = sum(more_extreme))
```

So 1159 times out of 10000 we observed a simulated test statistic less than -0.0776.
The p-value is 0.264. Note that we add one to the numerator to include the actually
observed difference of means of -0.0776.

$$
p-\mbox{Value} = \frac{1159+1}{10000} = 0.116.
$$




## Question 2:

**To be handed in on paper**: Say we have a population where the racial
breakdown of the juror pool (registered voters) is:

Race | Caucasian | African-American | Latino | Other | Total
----- | -----: | -----: | -----: | -----: | -----: | -----
Proportion | 72% | 7% | 12% | 9% | 100%

Social scientists want to know if there is a racial bias as to who gets selected
for jury duty in this population. They collect data on three scenarios of jury
selection, where each time 100 potential jurors are selected from the
population.

Scenario | Caucasian | African-American | Latino Count | Other Count | Total Count
----- | -----: | -----: | -----: | -----: | -----: | -----
A | 72 | 7 | 12 | 9 | 100
B | 0 | 0 | 100 | 0 | 100
C | 75 | 6 | 11 | 8 | 100

1. What is $H_0$ and $H_A$ for this study?
1. What is the test statistic for this study? Write it out in general.
1. What is the null distribution for this study? 
1. For each of these three scenarios individually, use the chi-squared test with $\alpha=0.1$ to answer the question
"is there racial bias in jury selection?" being sure to:
    a) Compute the *observed test statistic*
    a) Draw the null distribution and indicate the observed test statistic
    a) Compute the p-value. [Hint](https://beta.rstudioconnect.com/connect/#/apps/2719/access).
    a) Complete the test
    a) Write down your conclusion in non-statistical language


#### 1. Hypothesis Test Setup

* $H_0$: the selection was done randomly
* $H_A$: the selection was done with racial bias

Note: chi-square tests do **not** involve a population parameter


#### 2. Test Statistic

The test statistic in general is the **chi-squared test statistic**:

$$
\chi^2 = \sum_{i=1}^k \frac{(\mbox{Observed Count}_i - \mbox{Expected Count}_i)^2}{\mbox{Expected Count}_i}
$$
Note that when we observe exactly what we expect, the numertor is 0. So:

* Smaller values of $\chi^2$ indicates more evidence in favor of $H_0$
* Larger values of $\chi^2$ indicates more evidence in favor of $H_A$

In our case, we have $k=4$ groups; each of the 4 ethnic groups, so

$$
\chi^2 = \frac{(\mbox{Observed Caucasian Count} - \mbox{Expected Caucasian Count})^2}{\mbox{Expected Caucasian Count}} + \ldots + \frac{(\mbox{Observed Other Count} - \mbox{Expected Other Count})^2}{\mbox{Expected Other Count}}
$$

#### 3. Null Distribution

In general, the null distribution is not the normal distribution but rather
the $\chi^2$ distribution with $k-1$ degrees of freedom. Note: its values are
from 0 to $\infty$. In our case, since $k=4$, the degrees of freedom are $4-1=3$

```{r, echo=FALSE}
df <- 3
xlim <- qchisq(c(0, 0.99), df=df)
dchisq_limit <- function(x, obs_test_stat, df) {
  y <- dchisq(x, df=df)
  y[x < obs_test_stat] <- NA
  return(y)
}

base_plot <- ggplot(data = data_frame(x = xlim), aes(x)) +
  stat_function(fun = dchisq, n = 1000, args = list(df = df)) +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank())

base_plot + 
  labs(x="Chi-Squared Value", y="", title="Chi-Squared with df=3")
```

#### 4. Each Scenario

**Scenario 1:** In this case, the observed test statistic is 

```{r}
observed <- c(72, 7, 12, 9)
expected <- c(72, 7, 12, 9)
sum((observed-expected)^2/expected)
```

$$
\chi^2 = \frac{(72-72)^2}{72} + \frac{(7-7)^2}{7} + \frac{(12-12)^2}{12} + \frac{(9-9)^2}{9} = 0
$$

```{r, echo=FALSE}
obs_test_stat <- 0
area <- 1-pchisq(q=obs_test_stat, df=df) %>% 
  round(3)
title <- paste("Area to the right of red line is", area)

base_plot + 
  stat_function(fun = dchisq_limit, n = 1000, geom="area", fill="blue", alpha=0.2, args = list(obs_test_stat=obs_test_stat, df=df)) + 
  ylim(c(0, NA)) + 
  labs(x="Chi-Squared Value", y="", title=title) +
  geom_vline(xintercept = obs_test_stat, col="red")
```

Since the p-value is 1, for ANY $\alpha$ we'd have p-value $> \alpha$, so we do
not reject. Of course we do not reject the hypothesis of no racial bias; we
observed exactly what we'd expect!


**Scenario 2:** In this case, the observed test statistic is 

```{r}
observed <- c(0, 0, 100, 0)
expected <- c(72, 7, 12, 9)
sum((observed-expected)^2/expected)
```

$$
\chi^2 = \frac{(0-72)^2}{72} + \frac{(0-7)^2}{7} + \frac{(100-12)^2}{12} + \frac{(0-9)^2}{9} = 733.3333
$$


```{r, echo=FALSE}
obs_test_stat <- 733.3333
area <- 1-pchisq(q=obs_test_stat, df=df) %>% 
  round(3)
title <- paste("Area to the right of red line is", area)

base_plot + 
  stat_function(fun = dchisq_limit, n = 1000, geom="area", fill="blue", alpha=0.2, args = list(obs_test_stat=obs_test_stat, df=df)) + 
  ylim(c(0, NA)) + 
  labs(x="Chi-Squared Value", y="", title=title) +
  geom_vline(xintercept = obs_test_stat, col="red")
```

The p-value is 0; this is **very** extreme. So for ANY $\alpha$ we'd have p-value $< \alpha$, so 
we reject $H_0$ in favor of $H_A$. Of course we reject the hypothesis of no racial bias; how
could we have 100 out of 100 Latinos when they are only 12% of the population?


**Scenario 3:** Now things aren't as clear cut!

```{r}
observed <- c(75, 6, 11, 8)
expected <- c(72, 7, 12, 9)
sum((observed-expected)^2/expected)
```

$$
\chi^2 = \frac{(75-72)^2}{72} + \frac{(6-7)^2}{7} + \frac{(11-12)^2}{12} + \frac{(8-9)^2}{9} = 0.4623016
$$


```{r, echo=FALSE}
obs_test_stat <- 0.4623016
area <- 1-pchisq(q=obs_test_stat, df=df) %>% 
  round(3)
title <- paste("Area to the right of red line is", area)

base_plot + 
  stat_function(fun = dchisq_limit, n = 1000, geom="area", fill="blue", alpha=0.2, args = list(obs_test_stat=obs_test_stat, df=df)) + 
  ylim(c(0, NA)) + 
  labs(x="Chi-Squared Value", y="", title=title) +
  geom_vline(xintercept = obs_test_stat, col="red")
```

The p-value 0.927 is $> \alpha=0.1$, so we do not reject $H_0$. In other words
**while there still may exist racial discrimination, the data/evidence we
observed in Scenario 3 doesn't support this**.
