
---
title: "Problem Set 10"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, April 28, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
---

Load necessary packages:

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(readr)
set.seed(76)
```



## Question 1:

Answer the question "Did Econ majors do better on the intro stats final than
non-Econ majors?" using the `grades.csv` data set from Lec28. Note: save the
file `grades.csv` in the same folder/directory as `PS-10.Rmd`.


#### Hypothesis Test Setup

We have

* $H_0: \mu_E - \mu_O = 0$ vs
* $H_A: \mu_E - \mu_O > 0$

or equivalently

* $H_0: \mu_O - \mu_E = 0$ vs
* $H_A: \mu_O - \mu_E < 0$

Note the *one-sided* alternative of did Econ majors do better!


#### Observed Test Statistic

First, we compute the observed test statistic, noting carefully that it's the
mean of the other students minus the mean of the econ students: 
$\overline{x}_O - \overline{x}_E$:

```{r, message=FALSE}
grades <- read_csv("grades.csv")
mean(final ~ major, data=grades)
mean(final ~ major, data=grades) %>% diff()
observed_diff <- mean(final ~ major, data=grades) %>% diff()
```

It seems that Econ majors did 7.76% better, but is this difference **statistically** significant?

#### Null Distribution

Next, we construct the null distribution of the difference in means *assuming* there
is no true difference. Note when computing `difference`, we make sure that we
subtract in the correct order:

```{r, message=FALSE}
simulations <- do(1000) * mean(final ~ shuffle(major), data=grades)
simulations <- simulations %>%
  mutate(difference=Other-Economics)
```

#### Comparison

Finally, we compare the

* observed test statistic
* null distribution of the test statistic

```{r, message=FALSE}
ggplot(data=simulations , aes(x=difference)) +
  geom_histogram() +
  labs(x="Avg of Others - Avg of Economics") +
  geom_vline(xintercept = observed_diff, col="red")
```

The observed difference of -0.076 = -7.6% doesn't seem all that extreme. i.e.
it seems plausible 






## Question 2:

**To be handed in on paper**: Say we have a population where the racial
breakdown of the juror pool (registered voters) is:

Race | White | Black | Hispanic | Other | Total
----- | -----: | -----: | -----: | -----: | -----: | -----
Proportion | 72% | 7% | 12% | 9% | 100%

Social scientists want to know if there is a racial bias as to who gets selected
for jury duty in this population. They collect data on three scenarios of jury
selection, where each time 100 potential jurors are selected from the
population.

Scenario | White Count | Black Count | Hispanic Count | Other Count | Total Count
----- | -----: | -----: | -----: | -----: | -----: | -----
A | 72 | 7 | 12 | 9 | 100
B | 0 | 0 | 100 | 0 | 100
C | 75 | 6 | 11 | 8 | 100

1. What is $H_0$ and $H_A$ for this study?
1. What is the test statistic for this study? Write it out in general.
1. What is the null distribution for this study? 
1. For each of these three scenarios individually, use the chi-squared test with $\alpha=0.1$ to answer the question
"is there racial bias in jury selection?" being sure to:
    a) Compute the *observed test statistic*
    a) Draw the null distribution and indicate the observed test statistic
    a) Compute the p-value. [Hint](https://beta.rstudioconnect.com/connect/#/apps/2719/access).
    a) Complete the test
    a) Write down you conclusion in non-statistical language
