
---
title: "Problem Set 11"
author: "WRITE YOUR NAME HERE"
date: "Due Friday, May 5, 2017"
output:
  html_document:
    theme: cosmo
    highlight: tango
---

Load necessary packages:

```{r, message=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(mosaic)
library(fivethirtyeight)
data(congress_age)
set.seed(76)
```


## Scenario:

You are charged with studying the average age of members of congress (House of
Representatives and Senate) at the start of their term for all sessions of
Congress between:

* The [80th United States Congress](https://en.wikipedia.org/wiki/80th_United_States_Congress) 
from January 3, 1947 to January 3, 1949
* The [113th United States Congress](https://en.wikipedia.org/wiki/113th_United_States_Congress) 
from January 3, 2013 to January 3, 2015

In particular, you are interested in answering the following question:
comparing just the 80th and the 113th sessions, are members of congress on
average younger now (in 2013) or then (in 1947)?

However, researching birthdays and age at the start of the terms for all 18635 
members of congress between the 80th and 113th sessions of Congress is 
painstakingly boring work. So you decide to compute ages only for a random
sample of members of congress instead. 





## Question 1: CI for Difference of Population Means

Run these two lines to get your random sample of

* $n_{80}$ = 40 members of congress from the 80th session
* $n_{113}$ = 35 members of congress from the 113th session

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
congress_age_80 <- congress_age %>% 
  filter(congress == 80) %>% 
  sample_n(40)
congress_age_113 <- congress_age %>% 
  filter(congress == 113) %>% 
  sample_n(35)
congress_80_and_113 <- bind_rows(congress_age_80, congress_age_113)
```


#### a) EDA

Create one plot of exploratory data analysis. 

Note here we convert the variable `congress` from a numerical variable to a
categorical one by using `as.factor()`. It appears members of congress are older
now:

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
ggplot(data=congress_80_and_113, aes(x=as.factor(congress), y=age)) + 
  geom_boxplot()
```


#### b) 95% CI

Create a 95% confidence interval for $\mu_{80}-\mu_{113}$, the true **population
difference in means** of the average age in the 80th session minus the average 
age in the 113th session. **Hint**: The formula to approximate the standard
error for a difference of means $\overline{x}_1 - \overline{x}_2$ is

$$
SE_{\overline{x}_1 - \overline{x}_2} = \sqrt{ 
\left( \frac{(n_1-1)s_1^2 + (n_2-1)s_2^2}{n_1+n_2-2} \right) 
\left( \frac{1}{n_1} + \frac{1}{n_2} \right) 
}
$$

Note that the formula involves $s^2$ and not $s$, the sample variance = the sample standard deviation squared.

```{r, message=FALSE, echo=TRUE, warning=FALSE, fig.width=16/2, fig.height=9/2}
n1 <- 40
n2 <- 35
s1 <- sd(congress_age_80$age)
s2 <- sd(congress_age_113$age)
SE <- sqrt(  ((n1 -1)*s1^2 + (n2 -1)*s2^2)/(n1+n2-2) * (1/n1 + 1/n2)  )
SE
```

The point estimate $\overline{x}_{80} - \overline{x}_{113}$ based on our results
below is 51.54 - 55.82 = -4.28

```{r}
congress_80_and_113 %>% 
  group_by(congress) %>% 
  summarise(avg_age = mean(age))
```

So the 95% confidence interval is

```{r}
c(-4.28 - 2*SE, -4.28 + 2*SE)
```



**Your Answers**: 

* My confidence interval is: (-9.35, 0.79)
* Since the confidence interval includes 0, the data in the sample is not
suggestive of a meaningful difference in ages.
