---
title: "Problem Set Discussions"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: cosmo
    highlight: tango
    df_print: kable
---

<style>
h1{font-weight: bold;}
h2{color: #3399ff;}
h3{color: #3399ff;}
slides > slide.backdrop {background: white;}
</style>

```{r setup, include=FALSE}
# Ignore this:
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width=8, fig.height=4.5)
set.seed(76)
library(nycflights13)
library(dplyr)
library(ggplot2)
if(FALSE){
  knitr::purl(input="PS/PS.Rmd", output="PS/PS_code.R", documentation=2)
}
```




<!----------------------------------------------------------------------------->
# Problem Set 06

## Question 1: Drug Use Amongst OkCupid Users

Let's revisit the 60,000 San Francisco OkCupid users in 2012 and consider the 
variable `drug` reflecting users' answers to a question on their drug use. Run
this code first:

```{r, echo=TRUE}
library(okcupiddata)
data(profiles)
# Convert as missing values in drugs to an actual category "N/A"
profiles <- profiles %>% 
  mutate(drugs = ifelse(is.na(drugs), "N/A", drugs))
```


**a)** Type in a series of commands that will output a table of how many men and
women there are in this population.

**Solution:** This population skews male.

```{r, echo=TRUE}
# Write your code below:
male_vs_female <- profiles %>% 
  group_by(sex) %>% 
  summarise(count=n())
male_vs_female
```


**b)** Create a visualization that shows the distribution of the different
responses of in variable `drugs`.

**Solution:**

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar()
```

**c)** Create a visualization that shows the same information as in part a), but
for men and women separately. Who is more likely to say they never use drugs?
Men or women?

**Solution:** We plot the raw counts of drug use split by sex. However, we can't 
compare these values directly because the population involves 35,829 males and
24,117 females as we saw above!

```{r, echo=TRUE}
ggplot(data=profiles, aes(x=drugs)) +
  geom_bar() +
  facet_wrap(~sex)
```

**Advanced**: So instead we must normalize to proportions. Note this was not 
expected of you for this problem set at the time, but we are now in a position 
to understand it. Here we `group_by` both `sex` and `drugs`, get counts, and 
compute proportions, we change the `group_by()` structure, and then instead of
summarizing, we mutate to compute a proportion. Ask yourself, what values sum to
1?

```{r, echo=TRUE}
proportions <- profiles %>% 
  group_by(sex, drugs) %>% 
  summarise(count = n()) %>% 
  group_by(sex) %>% 
  mutate(proportion = count/sum(count)) %>% 
  mutate(proportion = round(proportion, 2))
proportions
```

Now let's try to plot it. Recall when we covered barplots, we saw there are two cases of barplots:

1. Where the counts are not pre-computed. what we've been using
1. Where the counts are pre-computed. This case.

We map the `proportion` to the `y` aesthetic. But for this to work, we need to
add a note to `geom_bar()` that we are specifying a `y` aesthetic by setting
`geom_bar(stat="identity")`

```{r, echo=TRUE}
ggplot(data=proportions, aes(x=drugs, y=proportion)) +
  geom_bar(stat="identity") +
  facet_wrap(~sex)
```

We now see that it is women who are more likely to self report that they never
use drugs.

**d)** After we loaded the data above via `data(profiles)`, we made sure to
convert all missing values, encoded in R as `NA`, to a specific response `N/A`
i.e. not answered. Why do you think it was important to do so?

**Solution:** Given the nature of the variable, some people might not be willing
to answer. It is important not to ignore these people or we might **bias** the
results.






## Question 2: Gapminder

We're going revisit the Gapminder data from PS-05. Run this code first:

```{r}
library(gapminder)
data(gapminder)
```


**a)** Output a table that shows the mean and median GDP of all countries in the
year 2007, but split by continent. Your table should be 5 rows and 3 columns.


```{r}
mean_median_GDP_by_continent <- gapminder %>% 
  filter(year==2007) %>% 
  select(continent, gdpPercap) %>%
  group_by(continent) %>% 
  summarise(
    mean_GDP_per_capita = mean(gdpPercap),
    median_GDP_per_capita = median(gdpPercap)
  )
mean_median_GDP_by_continent
```









## Question 3: Titanic

Let's study data relating to who survived the Titanic disaster. Run this code
first:

```{r, echo=TRUE}
data(Titanic)
# Convert the Titanic data to data frame format
Titanic <- Titanic %>% 
  tbl_df()
```


**a)** Output tables that compares survivor counts

1. between men and women
1. betweent the different classes


**Solution:** Let's first look at the `Titanic` data

```{r, echo=TRUE}
Titanic
```

We see that the variable `n` reports the **pre-computed** count of passengers
for each `Class`, `Sex`, `Age`, `Survived` category. The key difficulty of this
question was distiguising between using `n()` to count observations vs `sum()`.
Here, we need to `sum(n)`:

```{r, echo=TRUE}
survival_by_sex <- Titanic %>% 
  group_by(Sex, Survived) %>% 
  summarise(n=sum(n))
survival_by_sex

survival_by_class <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(n=sum(n))
survival_by_class
```



**b)** For each comparison above, indicate who was most likely to survive.

* `Survival` by `Sex`: Female survival 344/(344+126) = 73.2%, Male survival 367/(367+1362) = 21.2%. Women were much more likely to survive.
* `Survival` by `Class`: 1st class was most likely to survive at 62.5%

**Advanced**: convert to proportions

```{r, echo=TRUE}
survival_by_class_prop <- Titanic %>% 
  group_by(Class, Survived) %>% 
  summarise(n=sum(n)) %>% 
  group_by(Class) %>% 
  mutate(proportion = n/sum(n)) %>% 
  mutate(proportion = round(proportion, 3))
survival_by_class_prop
```














<!----------------------------------------------------------------------------->
# Problem Set 05

Load needed packages:

```{r}
library(ggplot2)
library(dplyr)
library(fivethirtyeight)
library(Quandl)
library(lubridate)
library(tidyr)
```

## Question 1: Bechdel Test

**a)** What is the Bechdel test? Hint: Read the first three paragraphs of the news
article linked in the help file for the bechdel data set.

**b)** The following bit of code takes the bechdel data set, limits to movies made 
in 1972 or later, and for each year reports the proportion of movies where 
binary is equal to PASS. Run this code (such code is the focus of Chapter 5: 
Data Wrangling):

```{r, cache=TRUE}
bechdel_cleaned <- bechdel %>% 
  select(year, binary) %>% 
  filter(year >= 1972) %>% 
  group_by(year) %>% 
  summarize(prop_pass_bechdel=sum(binary=="PASS")/n())
```

After `View()`'ing the data, create an appropriate plot that shows the time 
trend of the proportion of movies that pass the Bechdel test. Write your code
here:

**c)** What can you tell someone who asks "What is the state of female
representation" in movies?


### Discussion

**a)** For any of the data sets in the `fivethirtyeight` package, the help file 
lists information. For example: `?bechdel`. The Bechdel test is:

1. Are there at least two named women in the movie?
1. Do they have a conversation with each other at some point?
1. Does that conversation involve something other than a male character??

**b)** Notice how we computed the propotion using `group_by()`, `summarize()`,
and the `n()` summary/many-to-one function. Also note I forgot to erase the
solution to this question, so let me make the plot a little fancier:

```{r, cache=TRUE}
ggplot(bechdel_cleaned, aes(x=year, y=prop_pass_bechdel)) +
  geom_line() +
  labs(x="Year", y="Proportion Pass Bechdel Test", 
       "Proportion of Movies that Pass Bechdel Test over Time") +
  geom_smooth(se=FALSE)
```

`geom_smooth()` adds a **smoother** line, `se=FALSE` suppresses the error bars (
try `geom_smooth()` by itself where `se` is set to TRUE by default to see
difference). This picks out the **signal** from the **noise**.

**c)** Many possible answers here. In my opinion, there has been an increases 
since the 1970's in movies that pass Bechdel, but of late things seem to have 
leveled off at about 50%.



## Question 2: Cheese and Milk Production 

Run the code below. It will create a data frame food contains cheese and milk
production in the United States for every year since 1925. See these links for
more info:

* <https://www.quandl.com/data/USDANASS/NASS_CHEESEPRODUCTIONMEASUREDINLB>
* <https://www.quandl.com/data/NASS_MILKPRODUCTIONMEASUREDINLB>

```{r, cache=TRUE}
cheese <- 
  Quandl("USDANASS/NASS_CHEESEPRODUCTIONMEASUREDINLB", start_date="1925-01-01") %>% 
  mutate(type="cheese", date=ymd(Date), value=Value) %>%
  select(type, date, value) 
milk <- 
  Quandl("USDANASS/NASS_MILKPRODUCTIONMEASUREDINLB", start_date="1925-01-01") %>% 
  mutate(type="milk", date=ymd(Date), value=Value) %>%
  select(type, date, value) 
food <- bind_rows(cheese, milk) %>% 
  tbl_df()
```

**a)** Between cheese and milk, relatively speaking, which agricultural good has 
seen the bigger overall increases in production? Write your code below:


### Discussion

At first glance, it seems milk made the biggest gains:

```{r, cache=TRUE}
ggplot(data=food, aes(x=date, y=value, col=type)) + 
  geom_line()
```

But the above plot is of little use since the scale of milk prices is making it
difficult to see any differences in the price of cheese. Let's plot them
separately.

```{r, fig.height=3, fig.width=5, cache=TRUE}
# Write your code below:
ggplot(data=milk, aes(x=date, y=value)) + 
  geom_line()
ggplot(data=cheese, aes(x=date, y=value)) + 
  geom_line()
```

Both trend up, but which had the biggest relative increases i.e. which had the **highest percent increase**?

$$
\text{Percent difference} = \frac{\text{Price in 2015} - \text{Price in 1925}}{\text{Price in 1925}} \times 100 \%
$$


```{r, eval=FALSE, echo=FALSE}
# Ignore this:
data_frame(
  food_type=c("milk", "cheese"),
  price_1925 = c(90699000000, 501096000),
  price_2014 = c(206046000000, 11450117000 )
) %>% 
  mutate(
    `Absolute Change` = price_2014 - price_1925,
    `% Change` = round(`Absolute Change`/price_1925 * 100, 2)
  ) %>% 
  knitr::kable()
```

|Food Type |  1925|   2014| Absolute Change| % Change|
|:---------|-----------:|------------:|---------------:|--------:|
|milk      | 90,699,000,000| 206,046,000,000|    115,347,000,000|   127.18%|
|cheese    |   501,096,000|  11,450,117,000|     10,949,021,000|  2185.01%|

Cheese wins by a landslide. 2100% increase.



## Question 3: Histogram 

**a)** Recreate the histogram in Question 2 of the midterm using the following data:

```{r, cache=TRUE}
example <- data_frame(
  x = c(rep(3, 1), rep(4, 4), rep(5, 10), rep(6, 4), rep(7, 1), rep(3:7, each=4)),
  group = c(rep(1, 20), rep(2, 20))
)
```

### Solution

```{r, echo=TRUE, cache=TRUE}
ggplot(data=example, aes(x=x)) +
  geom_histogram(bins=5) +
  facet_wrap(~group)
```

We could've also done `geom_histogram(binwidth=1)`



## Question 4: Drinks 

Let's look at the data set drinks from the `fivethirtyeight` package.

```{r}
data(drinks)
```

**a)** Which 3 countries drink the most total alcohol according to this data?

**b)** The following bit of code uses the `gather()` function from the `tidyr`
package to convert the data to tidy format and then arranges it by country. Do 
Question 3.b) from the midterm using another `geom` besides `boxplot`.

```{r, cache=TRUE}
drinks_tidy <- drinks %>%
  gather(type, servings, -c(country, total_litres_of_pure_alcohol)) %>%
  arrange(country)

ggplot(data=drinks_tidy, aes(x=type, y=servings)) + 
  geom_boxplot()
```

**c)** What type of alcoholic beverage is the most consumed on earth?


### Discussion

**a)** The easiest way to do this is to use `View(drinks)` and play with the
sorting arrows: Belarus, then Lithuania, then Andorra.

**b)** We really want to emphasize the **beer vs spirits vs wine** comparison 
of the numerical variable `servings` above all:

```{r, cache=TRUE}
ggplot(data=drinks_tidy, aes(x=servings)) + 
  geom_histogram(binwidth = 50) +
  facet_wrap(~type)
```

**c)** In my opinion, this question is best answered using the boxplot: Overall
beer seems to be the most popular in terms of servings.



## BONUS

In the TED talk [The best stats you've ever
seen](https://www.ted.com/talks/hans_rosling_shows_the_best_stats_you_ve_ever_seen)
Hans Rosling (RIP 2017) presents human and international development data. The 
data seen in the video is accessible in the gapminder data set within the 
gapminder package:

```{r}
library(gapminder)
data(gapminder)
```

The next bit of code

* Filters the data set to only consider observations for two years: 1952 & 2007
* Renames the "pop" variable to "population"

```{r, cache=TRUE}
gapminder <- 
  filter(gapminder, year==1952 | year==2007) %>% 
  rename(population=pop)
```

**a)** Recreate the scatterplot of "Child Survival (%)" over "GDP per capita ($)"
seen at 12:00 in the video, but

* Making a comparison between 1952 and 2007
* Displaying "life expectancy" instead of "Child Survival"

Do so by changing the code snippet below:
```{r, eval=FALSE, echo=TRUE}
ggplot(data=DATASETNAME, aes(AES1=VAR1, AES2=VAR2, AES3=VAR3, AES4=VAR4)) +
  geom_point() + 
  facet_wrap(~VAR5) +
  scale_x_log10() + 
  labs(x="WRITE INFORMATIVE LABEL HERE", y="WRITE INFORMATIVE LABEL HERE", title="WRITE INFORMATIVE TITLE HERE")
```

**b)** Describe two facts that would be of interest to international development organizations.


### Discussion

**a)** 

```{r, cache=TRUE}
ggplot(data=gapminder, aes(x=gdpPercap, y=lifeExp, size=population, col=continent)) +
  geom_point() + 
  facet_wrap(~year) +
  scale_x_log10() + 
  labs(x="GDP per capita ($)", y="Life Expectancy", title="Life Expectancy over GDP per Capita")
```

**b)** In my opinion

* Overall, there have been gains in both GDP per capita and life expectancy
between 1952 and 2007
* However, broken down by continent:
    + Europe and the Americas have seen modest gains
    + Asia has made massive gains
    + Unfortunately Africa still lags behind










<!----------------------------------------------------------------------------->
# Problem Set 03

Load needed packages:

```{r}
library(ggplot2)
library(dplyr)
```


## Question 1: Movie Ratings

The `movies` data set in the `ggplot2movies` package has information and
ratings on 28,819 movies. This many data points is a bit unwieldy, so let's
take a random sample of 1000 of these movies. Furthermore, let's take the
variable `Comedy` and convert it to a `yes` vs `no` (binary) categorical
variable. Note: you don't need to understand this code for now, we'll see this
when we study data manipulation in Chapter 5.

```{r}
library(ggplot2movies)
data(movies)
movies <- movies %>%
  sample_n(1000) %>%
  mutate(Comedy=ifelse(Comedy==1, "yes", "no"))
```

a) You want to know for these 1000 randomly chosen movies: What is the
relationship between the year the movie was made and the IMDB rating?
Furthermore, I want to distinguish between comedies and non-comedies. In the
code block below, write the code that generates a graphic that will answer
this for you. Write your code here:


### Discussion

a) Comedy is a categorical variable we can split on by assigning a color to

```{r, cache=TRUE}
ggplot(data=movies, aes(x=year, y=rating, color=Comedy)) +
  geom_point()
```


b) This is very hard to eye-ball. We can add a **regression line** for Comedies
and Non-Comedies using `geom_smooth()`, which is a kind of `geom_line()`. We'll
see more and more of this as the semester progresses. Is that small difference
**statistically significant**? Hard to tell.


```{r, cache=TRUE}
ggplot(data=movies, aes(x=year, y=rating, col=Comedy)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```



## Question 2: Babynames

Considering the `babynames` data set in the `babynames` package again, we will
limit consideration to only the name "Casey".

```{r}
library(babynames)
data(babynames)
babynames <- babynames %>%
  filter(name=="Casey")
```

a) I want to know about popularity trends of the name "Casey" as a male name
and as a female name over the years. In the code block below, write the code
that generates a graphic that will answer this for you. Write your code here:

b) Given this graphic, what can you say about the name "Casey"? Don't merely
describe elements on graphic, but make a broader statement. For example, what
would you tell a parent who is interested in naming their child "Casey"?

### Discussion

a) While we could've made separate plots, its easier to view things on the same plot:

```{r, cache=TRUE}
ggplot(data=babynames, aes(x=year, y=prop, col=sex)) +
  geom_line()
```

b) This is a **unisex** name, which interestingly peaked in popularity at the same
time in the early 1980's. Take a look at this article from
[538](http://fivethirtyeight.com/features/there-are-922-unisex-names-in-america-is-yours-one-of-them/)
about unisex names.
