---
title: "Problem Set Discussions"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: cosmo
    highlight: tango
    df_print: paged
---

<style>
h1{font-weight: bold;}
h2{color: #3399ff;}
h3{color: #3399ff;}
slides > slide.backdrop {background: white;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width=8, fig.height=4.5)
set.seed(76)
library(nycflights13)
library(dplyr)
library(ggplot2)
```


<!----------------------------------------------------------------------------->
# Problem Set 03

Load needed packages:

```{r}
library(ggplot2)
library(dplyr)
```


## Question 1: Movie Ratings

The `movies` data set in the `ggplot2movies` package has information and
ratings on 28,819 movies. This many data points is a bit unwieldy, so let's
take a random sample of 1000 of these movies. Furthermore, let's take the
variable `Comedy` and convert it to a `yes` vs `no` (binary) categorical
variable. Note: you don't need to understand this code for now, we'll see this
when we study data manipulation in Chapter 5.

```{r}
library(ggplot2movies)
data(movies)
movies <- movies %>%
  sample_n(1000) %>%
  mutate(Comedy=ifelse(Comedy==1, "yes", "no"))
```

a) You want to know for these 1000 randomly chosen movies: What is the
relationship between the year the movie was made and the IMDB rating?
Furthermore, I want to distinguish between comedies and non-comedies. In the
code block below, write the code that generates a graphic that will answer
this for you. Write your code here:


### Discussion

a) Comedy is a categorical variable we can split on by assigning a color to

```{r}
ggplot(data=movies, aes(x=year, y=rating, color=Comedy)) +
  geom_point()
```


b) This is very hard to eye-ball. We can add a **regression line** for Comedies
and Non-Comedies using `geom_smooth()`, which is a kind of `geom_line()`. We'll
see more and more of this as the semester progresses. Is that small difference
**statistically significant**? Hard to tell.


```{r}
ggplot(data=movies, aes(x=year, y=rating, col=Comedy)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```



## Question 2: Babynames

Considering the `babynames` data set in the `babynames` package again, we will
limit consideration to only the name "Casey".

```{r}
library(babynames)
data(babynames)
babynames <- babynames %>%
  filter(name=="Casey")
```

a) I want to know about popularity trends of the name "Casey" as a male name
and as a female name over the years. In the code block below, write the code
that generates a graphic that will answer this for you. Write your code here:

b) Given this graphic, what can you say about the name "Casey"? Don't merely
describe elements on graphic, but make a broader statement. For example, what
would you tell a parent who is interested in naming their child "Casey"?

### Discussion

a) While we could've made separate plots, its easier to view things on the same plot:

```{r}
ggplot(data=babynames, aes(x=year, y=prop, col=sex)) +
  geom_line()
```

b) This is a **unisex** name, which interestingly peaked in popularity at the same
time in the early 1980's. Take a look at this article from
[538](http://fivethirtyeight.com/features/there-are-922-unisex-names-in-america-is-yours-one-of-them/)
about unisex names.
